---
title: "Module 2"
date: "March 9, 2018"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, size = "tiny",
                      fig.height = 2.5,
                      fig.width = 4.5,
                      fig.align = "center", width = 50, 
                      tidy.opts=list(width.cutoff=48),tidy=TRUE)

#install.packages("stringr")
library(stringr)
library(spatstat)
library(dplyr)
library(scales)
library(ggplot2)
library(stargazer)
library(Hmisc)
library(lubridate)
library(readr)
#install.packages("scales")
library(scales)
setwd("/if/fame/afe/usr/amonken/Howard/Class_Materials/Spring_2018/module_2/day_2")
```

# Day 2

## Recap Last Week

\centerline{\includegraphics[height=3in]{./Reference/for_loop.png}}

## For-loops: Warm Up Exercise

* Cubing every number from 1 to 8 and print the results

```{r, eval = F}

for(num in    ){
    print(     **3) 
}

```

## Recap Last Week - In-Class Exercise
* Like we did last class, let's convert this vector of substrings into a single string. Make sure to use meaningful variable names in your code. (Hint: Use the paste function.)

```{r, eval = FALSE}
string_vec <- c("Economics","is","the","best","subject!")

## initialize the variable

for(item in string_vec){
  ## code to be executed
}

```


```{r "In-class answer", include = FALSE}
string_vec <- c("Economics","is","the","best","subject!")

## initialize the variable
phrase = NULL

for(item in string_vec){
  phrase = paste(phrase, item)
}

print(phrase)
```


## Goals for Today

Economics - Question of the Day

* How does the sale price of residential property change based on various factors?

### What things will affect the property value?

## Goals for Today

* Determinants of property values
    + Size of house (square footage)
    + Size of property
    + Higher floors vs lower floors
    + Quality of local schools
    + Number of bedrooms and bathrooms
    + Neighborhood location (metro, restaurants, library)
    
## Goals for Today

Programming - R

* Reviewing for-loops and if/else Statements
* Introduction to stringr
* Combining data using `bindrows`, `%in%`, and `join` functions
* Cleaning data
* Effective graphics for data visualization

## If/Else Statements: Introduction
* We have previously used `ifelse()` to make decisions about recoding our variables with `mutate`
* If/else statements work in a similar way
* In R, a basic if/else statement takes the following form:
```{r, eval=FALSE}
if(logical argument){
  ## code to be executed
} else{
  ## code to be executed
}
```

## If/Else Statements: Example

\centerline{\includegraphics[height=3in]{./Reference/if_else_statement.png}}

## Evens and Odds Example
* We will use the modulo operator (`%%`) to characterize numbers as even or odd
* `%%` returns the remainder after division of the first argument by the second argument

\scriptsize
```{r, collapse=TRUE}
# %% is the modulus operator 
# We are finding the remainder!
4 %% 1; 10 %% 4; 3 %% 2
testNum <- 12
if(testNum %% 2 == 0){
  print(paste(testNum, "is EVEN"))
} else{
  # If the remainder after division by 2 is not 0 
  # then it must be odd (right..?)
  print(paste(testNum, "is ODD"))
}
```

## If/Else Statements Combined with For-Loops
* Let's characterize and record the numbers from 1 to 10 as even or odd

\scriptsize
```{r}
evens <- numeric()
odds <- numeric()

for(i in seq(1, 10, by = 1)){
  if(i %% 2 == 0){ # %% is the modulus operator 
                   # --- we are finding the remainder!
    evens <- c(evens, i)
  } else{
    odds <- c(odds, i)
  } }

evens
odds
```

## In-Class Exercise: If/Else Statements
* Given the following grade, use an if/else statement to determine if the student passed or failed
    + **(A passing grade is greater than or equal to 60)**
```{r}
student_grade <- 71
```
**Return a correct statement that the student passed or failed**

```{r, , echo = T, eval = F}
if(             ){
  print("     ")
} else {
  print("     ")
}
```

## Else-if statements: Introduction
* We aren't limited to choosing between two conditions
* Similar to `case_when()`, else-if lets us make multiple decisions
    + provides us with even more flexibility
* Checks each condition one by one
    + check the first condition, if false then it moves on to the the next one
* **Else catches everything that does not meet the previous criteria** so be careful when coding or deciding what to include

## Else-if statements: Example
* Here's an example of assigning survey numerical responses to genders.

\scriptsize
```{r}
response_list <- c(0,1,2,2,0,1,2,1,0,2,2,1,1,1,2,1,0)
female <- 0; male <- 0; other <- 0

for(person in response_list){
  if(person == 1) {
    male <- male + 1
  } else if (person == 2){
    female <- female + 1
  } else {
    other <- other + 1
  }
}
survey_genders <- data.frame(male,female,other)
survey_genders
```

## Else-if statements: Assigning Grades - In-Class Exercise

* You've been given a list of test scores that you want to categorize into letter grades
* Use if else statements to assign letters to the numeric grades
    + A 90 - 100, B 80 - 89, C 70 - 79, D 60 - 69, F < 60

\scriptsize
```{r, eval=FALSE}
test_scores <- c(85,55,100,67,73,92,94,99,87,89.3)

# Initialize our vector
letter_grades <-   

for(grade in test_scores){
  if(   >= 90){
    letter_grades <- paste(letter_grades,"A")
  } else if(   ) {
    letter_grades <-    (            ,"B")
  } else if
    letter_grades <- 
    else if 
    
  } else {
    letter_grades <- paste(       "F")
  }
}
letter_grades
```


## Reading in Data: Introduction

* Before we can access our datasets we need to help R find them
* We can create a vector of all the files in our data folder using the `list.files` function
* **What sorts of files are in the folder?**

\scriptsize
```{r}
## create a vector of dataset names
temp_files <- list.files("./Data/")
temp_files
```

## Reading in Data: Introduction

\small

* We need to be able to access these files, we'll need to tell R to look in the folder 'Data'
      + Therefore, we should append 'data' onto the title using `paste0`
      
\scriptsize
```{r}
files <- paste0("./Data/", temp_files)
files
```

\small

* Now that we know the names of the files with the datasets, we need to decide which datasets are useful and which go together
*  **What are the two major categories of files?**
*  **How can we combine these files most effectively?**

## Introduction to str_detect()
* `str_detect()` is a function in the tidyverse package that can tell whether a string contains a certain word or phrase. The function returns:
    + `TRUE` if the word you're looking for is in the string
    + `FALSE` if the word you're looking for is *not* in the string
    
\footnotesize
```{r}
str_detect("property_redfin_01.csv", "property")
str_detect("location_redfin_01.csv", "property")

str_detect(c("property_redfin_01.csv", 
             "location_redfin_01.csv"),
           "location")
```

## Using str_detect
* Using what we've learned about `str_detect`, let's create a vector that only includes location files.

\scriptsize
```{r}
files[str_detect(files, "location")]
```


## Building our Dataset - In-Class Exercise

\small

* You will need to use for-loops and if-else statements to create two datasets, one dataset that includes all the 'location' data and one dataset that includes all the 'property' data. 
    + Initialize two data frames to store all your new data: one data frame for the 'property data' and one for the 'location data'
    + Create a for-loop that cycles through each csv in the folder and checks if it is a 'location' or 'property file'.
    + If it's a property file, add (bind) the csv's data onto your property data frame
    + If it's a location file, bind it into your location_data data frame. 
**Hint(check out what the `bind_rows` function does.)**

\tiny
```{r, eval=FALSE}
property_data <- data_frame() # initialize empty data frame for the property data
 <- data_frame() # initialize empty data frame for the location data
for(file in        ){
    # You only want to have property data in the first dataset
    if(str_detect(     ,"property")){ 
         <- read_csv(   )
        property_data <- bind_rows(property_data, )
    } 
  # Now let's combine location data into a different dataset
  else if(str_detect(      , "location")){ 
    }
}
```


```{r, include = FALSE}
property_data <- data_frame() # initialize empty data frame
location_data <- data_frame() # initialize empty data frame

for(file in files){
# this is a property file, add to property_data
    if(str_detect(file, "property")){ 
        data <- read_csv(file)
        property_data <- bind_rows(property_data, data)
    } 
  # location file
  else if(str_detect(file, "location")){ 
        data <- read_csv(file)
        location_data <- bind_rows(location_data, data)
    }
}
```


## Overview of the Data

* Let's take a look at the variables in our two dataframes - property data and location data
* **What does the property data do? What does the location data do?**
* The good news is that we have much of the data we need, the bad news is that the data is split between location information and price information
* **What piece of information is available in both datasets?**
* We want to combine these two datasets around that common piece of information

## Comparing our Datasets

* Harnessing the power of the `%in%` function

\scriptsize
```{r}
vec1 <- c("green", "yellow", "dog", "teal", "violet", "beige", "cat", "red", "horse", "orange")
vec2 <- c("yellow", "red", "blue", "purple", "green", "orange")
# What does the %in% function show us?
vec1 %in% vec2

# Where are the words that overlap across vectors?
# Where are the words in vector 1 that appear in vector 2?
which(vec1 %in% vec2) 

# How do we return the names of the elements that appear in both vectors?
vec1[vec1 %in% vec2] 
```

## Where to Combine the Data

* Using our `%in%` function, let's find what variable(s) occur(s) in **both** our property and location datasets.
* We'll save the result as 'overlap', which is a variable we can use for our next step.

\scriptsize
```{r, echo = TRUE}
property_names <- names(property_data)
location_names <- names(location_data)

# What variable exists in both datasets?
property_names %in% location_names

overlap <- property_names[property_names %in% location_names]
overlap


```

## "Joining" the Data: Introduction

\small

* Our URL column exists in both of the dataframes!
* We only want houses that exist in both datasets
* We call combining data frames "joining." Below are types of joins:
* Which type of joining combines our dataframes so we only have properties with both location and property information?

\centerline{\includegraphics[height=2.4in]{./Reference/joins.png}}




## "Joining" the Data: In Practice

\footnotesize
```{r}
joined_data <- inner_join(property_data,
                         location_data,
                         by = overlap)

names(joined_data)
```

## "Joining" the Data: Additional Notes

* Success! Now we have one dataset with all of our housing data
* Note - It is also possible to perform a `join` where the columns we are matching in each table do not have the same name using the 'by' argument: by = c("name1" = "name2").
* Similarly, it is also possible to `join` based on multiple columns: by = c(column1, column2, etc...)

## Examining the Data

* Now we can examine the data to check for any problems.
* **What are some problems you see in terms of usability for this data?**
```{r include = F, eval = F}
View(joined_data)
```

## Examining the Data: Potential Problems

* price column - USD in front of the numbers
* propertytype column - has "propertytype" in front of every observation
* city_state column - state is inconsistent (va, VA, Virginia) and unwanted states
* lat_lon column - & in the middle of the string
* zip_code column - extra zeros
* sold.dates, list.date columns - dates are not consistent

## Cleaning the Data: Why We Do It
* If we try to make a plot without cleaning up our data...it doesn't work too well

```{r echo= FALSE, fig.width=4, fig.height=2}
joined_data %>% 
    ggplot(aes(x = SQUARE.FEET, y = PRICE)) +
    geom_point()
```

## Cleaning the Data: Price Column
* Let's start by looking at our price data more closely

\scriptsize
```{r}
head(joined_data$PRICE)
```

\normalsize

* There's a USD in front of each number of the column
* What class of variable is the price column?

## Cleaning the Data: Price Column

* How can we fix this problem for a character variable? What type of method could we use?

## Cleaning the Data: Price Column

* One way we could eliminate the USD is by replacing the USD with nothing, "" (this would be the same as removing it).
* Try using the str_replace function to modify the string "USD40000" to be "40000"

```{r eval = T, echo = T}
str_replace("USD40000", "USD", "")
```

## Cleaning the Data: Price Column

* Let's mutate the price column to remove "USD" from the character string

\scriptsize
```{r eval = T, echo = T, results = F}
joined_data <- joined_data %>% 
    mutate(PRICE = str_replace(PRICE, "USD", ""))
head(joined_data$PRICE)
```

## Cleaning the Data: A Second Attempt at Plotting

* Let's redo the plot!
```{r, echo= FALSE, fig.width=4.5, fig.height=2.5}
joined_data %>% 
    ggplot(aes(x = SQUARE.FEET, y = PRICE)) +
    geom_point() +
    labs(title = "Relationship between price and home size",
         y = "USD",
         x = "Sq. Ft.",
         caption = "Data from Redfin")
```

## Cleaning the Data: Reclassifying Columns and Removing N/As - In-Class Exercise
* We still have a problem, we forgot to convert our PRICE column to be class numeric!
* We also currently have a warning about missing values -- what should we do about those?
* You should: 
    + Convert the price column to be numeric
    + Remove any missing values from the price and square feet columns
    + Then replot and see how it looks
  
## Cleaning the Data: Reclassifying Columns and Removing N/As - In-Class Exercise

\scriptsize
```{r, eval= FALSE, fig.width=4.5, fig.height=2.5}
joined_data <- joined_data %>% 
    filter(!is.na(      ), !is.na(         )) %>%
    mutate(PRICE = as.numeric(     ))

# Now make the plot
joined_data %>% 
    ggplot(aes(x =      , y =      )) +
    geom_point() +
    #Note that the scale_y_continuous makes sure we don't get expotentials!
    scale_y_continuous("USD", labels = dollar) + 
    labs(title = 
           "Price vs. Home Size",
         y = ,
         x = ,
         caption = "Data from Redfin")
```
    
\normalsize

* What is the relationship between Sq Ft and home price?

```{r, include = FALSE}
joined_data <- joined_data %>% 
    filter(!is.na(PRICE), !is.na(SQUARE.FEET)) %>%
    mutate(PRICE = as.numeric(PRICE))

```




## Understanding the Data: Property Type and Price

* What about the relationship between property type (apt, single family house, condo, etc.) and price?
* Let's produce a table and a bar plot of the average price per square foot by property type.


## Cleaning the Data: Property Type - In-Class Exercise

* Let's clean the property type column
* Check out the PROPERTY.TYPE column and use `str_replace` to fix it

```{r, eval = F}
joined_data <-         %>% 
    mutate(PROPERTY.TYPE = str_replace(          ,
    "propertytype:",   ))
```


\scriptsize
```{r, include = FALSE}
joined_data <- joined_data %>% 
    mutate(PROPERTY.TYPE = str_replace(
      PROPERTY.TYPE, "propertytype:", ""))
```


## Understanding the Data: Property Price/Sq Ft Table
* Let's create a table of average price per square foot by property type

\scriptsize
```{r}
type_price <- joined_data %>% 
    group_by(PROPERTY.TYPE) %>% 
    summarise(price = mean(PRICE/SQUARE.FEET, na.rm = T))

type_price
```

## Understanding the Data: Property Price/Sq Ft Graph
* Now let's check out a graph. This is obviously not a production-quality graph, just something for our reference.

* How are the colors defined in the graph?

```{r, echo = FALSE}
type_price %>% 
            ggplot(aes(x = PROPERTY.TYPE, 
                       y = price, 
                       fill = PROPERTY.TYPE)) +
            geom_bar(stat = "identity") +
            scale_y_continuous("USD", labels = dollar) +
            scale_fill_discrete(guide = F) + # get rid of the fill legend
            theme(axis.text.x = element_text(angle = 20, hjust= 1, vjust = 1))
```

## Understanding the Data: Data Challenges

* City/state is another variable that might impact home sale price... Why?

```{r}
unique(joined_data$CITY_STATE)
```
* What issues do we have with the city_state variable?

## Cleaning the Data: String Split

* We can use the `str_split_fixed` function to alter our string (or vector of strings) using a 'pattern'
* The function returns the peices of the string after splitting based on your 'pattern', which could be a word, symbol, or number of characters.

\scriptsize
```{r}
# We have three arguments, what do they do?
strings <- str_split_fixed("United States of America", "of", n = 2)
strings

class(strings)
```

## Cleaning the Data: String Split

* We told R to split into two pieces (n = 2), so we got two columns in return.
* If we want to get back to individual elements, we can subset the matrix using brackets.

```{r}
strings[1, 1] # first row, first column
strings[1, 2] # first row, second column
```

## Cleaning the Data: String Split

* Let's test it out with something closer to our actual use case.
* Note that we're splitting along the comma and a space, which will remove both!

```{r}
str_split_fixed("Arlington, Virginia", ", ", n = 2)
```

## Cleaning the Data: String Split - In-Class Exercise

* Use what we just learned to mutate the city_state column and create separate city and state columns.

\scriptsize
```{r, eval = FALSE}
joined_data <-       %>% 
    mutate(CITY = str_split_fixed(CITY_STATE,   )[1, 1],
           STATE =     (       )[1, 2])

joined_data %>% select(CITY_STATE, CITY, STATE) %>% head
```


```{r, include = FALSE}
joined_data <- joined_data %>% 
    mutate(CITY = str_split_fixed(CITY_STATE, ", ", n = 2)[,1],
           STATE = str_split_fixed(CITY_STATE, ", ", n = 2)[, 2])

joined_data %>% select(CITY_STATE, CITY, STATE) %>% head
```

## Cleaning the Data: State Data

* We're not done yet. Let's take another look at the state column and see what types of responses exist

\scriptsize
```{r}
unique(joined_data$STATE)
```

\normalsize
* What do you think we need to do to further clean this column?

## Cleaning the Data: Capitalizing States

* The function `str_to_upper` will convert all lower case letter to upper case ones

\scriptsize
```{r}
str_to_upper("va")
```

* We'll use this function to mutate the state column to upper case

\scriptsize
```{r}
joined_data <- joined_data %>% 
    mutate(STATE = str_to_upper(STATE))

unique(joined_data$STATE)
```

## Cleaning the Data: Abbreviating States - In-Class Exercise
* Now we want to change "VIRGINIA" to "VA" in the column to make the label consistent
* We've already used this function! Go back and find the function you need to make that change

\scriptsize
```{r, eval = F}
joined_data <- joined_data %>% 
    mutate(STATE =            (     , "VIRGINIA",    ))
```

\scriptsize
```{r "VIRGINIA ANSWER", include = FALSE}
joined_data <- joined_data %>% 
    mutate(STATE = str_replace(STATE, "VIRGINIA", "VA"))
```


## Cleaning the Data: Removing Additional States

* The last thing we need to do is toss out the rows with MICHIGAN and COLORADO since we only want DMV data
* Use the `%in%` function to filter only rows with VA, DC, or MD in them.

\scriptsize
```{r, collapse = TRUE}
joined_data <- joined_data %>% 
    filter(STATE %in% c("VA", "DC", "MD"))

unique(joined_data$STATE)
```

## Plotting the Data: Price by State
* Let's find out the average prices by state for our dataset

\scriptsize
```{r, eval = FALSE}
state_average <- joined_data %>% group_by(STATE) %>% 
    summarise(price = mean(PRICE/SQUARE.FEET, na.rm = T))

state_average_plot <- state_average %>% 
    ggplot(aes(x = STATE, y = price, fill = STATE)) +
    geom_bar(stat = "identity", width = 0.5) + 
    labs(x = NULL, # What does this do?
         y = "USD/Sq ft",
         title = "Average price for DC Home sales",
         caption = "Data from Redfin") 

state_average_plot
```

## Plotting the Data: Price by State
* I want to have a larger plot title and have it centered
* How can we do that? Would we use a data-related dimension or an aesthetic option?

\scriptsize
```{r, echo = FALSE}
state_average <- joined_data %>% group_by(STATE) %>% 
    summarise(price = mean(PRICE/SQUARE.FEET, na.rm = T))

state_average_plot <- state_average %>% 
    ggplot(aes(x = STATE, y = price, fill = STATE)) +
    geom_bar(stat = "identity", width = 0.5) + 
    labs(x = NULL, # What does this do?
         y = "USD/Sq ft",
         title = "Average price for DC Home sales",
         caption = "Data from Redfin") 

state_average_plot
```

## Plotting the Data: Price by State
* We can use the `theme()` function to alter elements of the graph.
* What other textual elements can we alter in this graph?

\scriptsize
```{r, fig.width=4, fig.height=2}
state_average_plot +
    theme(plot.title = 
            element_text(size = 14, hjust = .5, color = "Red"))
```

## Plotting the Data: Appearance Options

* We can use `element_text` to control any text on our graph, including the title, axis labels, legend, etc. `element_text` allows us to set:

    + font family (Times New Roman, Arial, etc...) 
    + font size (10, 12, 14, etc...)  
    + font face (bold, italic) 
    + color 
    + hjust (horizontal adjustment) 
    + vjust (vertical adjustment)
    + angle  
    + other aspects that impact the display of text.

## Plotting the Data: Property Type

\scriptsize
```{r, , eval = FALSE, fig.width=5, fig.height=3}
state_proptype <- joined_data %>% 
    group_by(PROPERTY.TYPE, STATE) %>% 
    summarise(price = mean(PRICE/SQUARE.FEET, na.rm = T)

plot <- state_proptype %>% 
    ggplot(aes(x = STATE, y = price, fill = STATE)) +
    geom_bar(stat = "identity") +
    facet_wrap("PROPERTY.TYPE") +
    labs(title = "Price by property type",
         x = NULL,
         y = "USD/Sq ft.",
         caption = "Data from Redfin")

plot + theme(plot.title =
          element_text(face = "italic", size = 14, color = "Blue"),
          axis.text.x = element_text(face = "bold", color = "green4"),
          strip.text.x = element_text(size = 6, color = "black"),
          plot.background = element_rect(fill = "beige", color = "red"),
          panel.background = element_rect(fill = "cadetblue1", 
                                          color = "black", linetype = 3))
              
              
```

## Plotting the Data: Property Type
\normalsize
```{r, echo = FALSE, fig.width=4.5, fig.height=3}
state_proptype <- joined_data %>% 
    group_by(PROPERTY.TYPE, STATE) %>% 
    summarise(price = mean(PRICE/SQUARE.FEET, na.rm = T)) 
state_proptype %>% 
    ggplot(aes(x = STATE, y = price, fill = STATE)) +
    geom_bar(stat = "identity") +
    facet_wrap("PROPERTY.TYPE") +
    labs(title = "Price by property type",
         x = NULL,
         y = "USD/Sq ft.",
         caption = "Data from Redfin") +
    theme(plot.title = element_text(face = "italic", size = 14, color = "Blue"),
          axis.text.x = element_text(face = "bold", color = "green4"),
          strip.text.x = element_text(size = 6, color = "black"),
          plot.background = element_rect(fill = "beige", color = "red"),
          panel.background = element_rect(fill = "cadetblue1", 
                                          color = "black", linetype = 3))
              
              
```

## Understanding the Graphic Appearance

* We used `element_rect()` to adjust rectangular elements of the plot.
* The main plot has a beige background with a red border.
* We used the panel.background option to adjust elements of the area where the data is
* What did the linetype argument do in our panel.background option?

\scriptsize
```{r, eval = FALSE, fig.width=5, fig.height=3}
plot <- state_proptype %>% 
    ggplot(aes(x = STATE, y = price, fill = STATE)) +
    geom_bar(stat = "identity") +
    facet_wrap("PROPERTY.TYPE") +
    labs(title = "Price by property type",
         x = NULL,
         y = "USD/Sq ft.",
         caption = "Data from Redfin")

plot + theme(plot.title = element_text(face = "italic", size = 14, color = "Blue"),
          axis.text.x = element_text(face = "bold", color = "green4"),
          strip.text.x = element_text(size = 6, color = "black"),
          plot.background = element_rect(fill = "beige", color = "red"),
          panel.background = element_rect(fill = "cadetblue1", 
                                          color = "black", linetype = 3))
              
              
```

## Understanding the Data: Price per Lot Size

* Perhaps using the price per square foot is not the best method of analysis.
* When a property is sold the surrounding land around the house will likely affect the selling price.
* Let's work on making a plot of price/lot size

## Understanding the Data: Price per Lot Size

* We'll start by looking at the head of the data
```{r, echo = FALSE}
joined_data %>% select(PRICE, SQUARE.FEET, LOT.SIZE) %>% head()
```
* We have a lot of NA values. Why is that?

## Understanding the Data: Price per Lot Size - In-Class Exercise

\small

* Let's get rid of the `NA` values by replacing them with square footage
* We'll create a new column called Property Size that equals either lot size or square feet
      + Use an `ifelse()` statement in our mutate, where we test whether lot size is `NA`
      + If lot size is `NA`, use square feet. Otherwise, use lot size
* Next we will summarize our data by price per foot of property size. 
      + We want to group by property type and state!

\scriptsize
```{r, eval = FALSE}
joined_data <-         %>% 
    mutate(PROPERTY.SIZE = ifelse(is.na(       ), SQUARE.FEET,        ))

# Now we can find our average
lot_state_data <- joined_data %>% 
    group_by(          ) %>% 
    summarise(price = mean(                 , na.rm = T))

head(lot_state_data)
```

\scriptsize
```{r, include = FALSE}
joined_data <- joined_data %>% 
    mutate(PROPERTY.SIZE = ifelse(is.na(LOT.SIZE), SQUARE.FEET, LOT.SIZE))

# Now we can find our average
lot_state_data <- joined_data %>% 
    group_by(PROPERTY.TYPE, STATE) %>% 
    summarise(price = mean(PRICE/PROPERTY.SIZE, na.rm = T))

head(lot_state_data)
```


## Understanding the Data: Culminating Exercise
* Make a plot like the one we just did but using our new price/property.size variable with some modifications -
    + **Make your title centered (using hjust) and green**
    + **Make your panel background white with a red border**

```{r "Lot size plot answer", eval = F, include = F}
lot_state_data %>% 
    ggplot(aes(x =      , y =     , fill =     )) +
    facet_wrap("PROPERTY.TYPE") +
    geom_bar(stat = "identity", width = 0.5) +
    labs(x = NULL,
         title = "      ",
         caption = "           ",
         y = "         ") +
    theme(plot.title = element_text(hjust = , color =        ),
          strip.text.x = element_text(size = 6, color = "black"),
          panel.background = element_rect(             ))
```