---
title: "Module 3"
output: beamer_presentation
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, size = "tiny", width = 50, fig.height = 3.5, fig.width = 5)

library(tidyverse)
library(plotly)
library(gapminder)
library(maps)
library(stargazer)
library(ggthemes)
```

# Day 2: Maps!!!

## Recap of Last Week

* Fixed Effects 

* Differences-in-Differences 

## Recap: Fixed Effects 

* Fixed effects are a tool that allow us to control for unobserved variation through the use of dummy variables 

* A \textbf{Factor Model} is an econometric model that uses fixed effects to control for the variation for the variables that are not of interest

* Pros
    * Can correct for omitted variable bias
    * Low computational cost

* Cons 
    * Need large amounts of data for your variables of interest to keep statistical significance  
    * Might lead to over fitting 
  
## Recap: Fixed Effects pt. 2

Suppose we have annual wage data for 20,000 individuals that spans from 2000-2017. We want to investigate what effect does education have on wages, so we estimate the following model 

$$ WAGES_{it} = \beta_0 + \beta_1EDUC_{it} + \epsilon_{it}$$
Where $WAGES_{it}$ is the annual wages and $EDUC_{it}$ is the years of education individual $i$ has in time $t$

* Is this Panel, Cross Sectional, or Time Series data?

* What are some ommited variables? 

## Recap: Fixed Effects pt. 3

We can control for all of those omitted variables by using individual and fixed effects!

$$ WAGES_{it} = \beta_0 + \beta_1EDUC_{it} + \delta_i + \lambda_t + \epsilon_{it}$$
Where $\delta_i$ is a dummy variable for each individual!

* How many total dummy variable are there? 

* What happens if we include time fixed effects (dummy variables for each year)? 

## Recap: Differences-in-Differences 

* The ingredients
    * Treatment - policy intervention
    * Control group
    * Treated group
  
* Look at the difference between the marginal effect between the treated and control group to back out the impact of the treatment 

## Recap: Differences-in-Differences pt. 2

* Pros
    * Estiamtes the causal effect of the treatment in a rigirous and well understood manner 
    * Easy and fast to estimate
* Cons 
    * Very specific type of data 
    * Evidence of parallel trends 
    * Only describes the effect of treatment not how the mechanism works 
  
## Recap: Differences-in-Differences pt.3 

* Last time we recreated Card and Kruger (AER 1994)

* The authors investiagted if an increase in the mimum wage leads to a decrease in employment

* The ingrediants 
    * Treatment - law that increased minimum wages
    * Control group - New Jersey
    * Treated group - Pennsylvania

* They find that there was no effect of the minumum wage increase on employment! 

* Any thoughts? What did they do well? What could have been improved? 
  
## Recap: Differences-in-Differences pt.4

```{r, height = 2, width = 5, echo=FALSE}
didData <- data_frame(Year = rep(2000:2004, 2),
                      Treatment = c(rep("Treated", 5), 
                      rep("Control",5)),
                      Y = c(3:5, 7, 9, 1:5))

counterfac <- data_frame(Year = rep(2000:2004, 2),
                         Y = rep(3:7, 2))

ggplot(didData, aes(x = Year, y = Y, colour = as.factor(Treatment))) +
  geom_line(lwd = 1.5) +
  labs(x = NULL, y = "Outcome", colour = "Treatment", title = "Example of Parallel Trends") +
  geom_line(aes(x = counterfac$Year, y = counterfac$Y), linetype = 4, lwd = 0.5) +
  geom_vline(aes(xintercept = 2002), linetype = 5) +
  theme_hc()

#Tidy up workspace
rm(didData)
rm(counterfac)
```



## Game Plan for the Day 

* Investigate some development data 
    * Measuring growth vs. measuring development
    * Precanned objects in functions 
    * Relationship between wealth and development 

* Plotting maps in R
    * Create maps with ggplot
    * Incoporate data with maps 
  
## Packages with Data Frames in Them

  * Some packages include sample data frames to play around with when you install it
  
  * Examples include Starwars Movie characters, stock prices, and movie ratings 
  
  * Precanned data frames allow you skip loading data in! 
  
## Loading in Data 
```{r}
library(gapminder)
world_data <- gapminder
head(world_data)
```

## Quick Glance at the data 

* What frequency is data reported at?

* What variable is measuring quality of life (i.e Development)?


## Development vs. Growth 

* \textbf{Economic Growth} is the study of how economies grow through technology, research, and investments innovations.

* What recent technological innovations do you think that have attributed to growth in the American economy?

* \textbf{Development Economics} is the study of how nations improve the economic, political, and social well-being of its people. 

* How can we measure development?

## The Relationship between Economic Growth and Development 

* Is there a correlation between a country's wealth and how "developed" it is?  

* How much of a country's development is due to the country's wealth?

* This is our research question for the day! 

* How would we investigate this relationship with our data? 

## Plotting the relationship between GDP Per Capita and Life Expectancy

```{r, echo = FALSE}
world_data %>%
  ggplot(aes(gdpPercap, lifeExp)) +
  geom_point() 
```



## In Class Exercise 1: Make a pretty plot!!!

* Y'all are now plotting pros! 

* Fix up the previous plot and differentiate the observations by contitinet. 

* I would suggest to use scale_x_log10(), if you do not know what this function does please google it or type ?scale_x_log10() into you counsle 

## In Class Exercise 1: Solution 

```{r, echo = FALSE}
world_data %>%
  ggplot(aes(gdpPercap, lifeExp, color = continent)) +
  geom_point() +
  scale_x_log10() +
  xlab("Log of GDP Per Capita") +
  ylab("Life Expectancey") +
  ggtitle("Relationship of GDP Per Capita and Life Expectancy")
  
```

## Creating a Benchmark Model 

```{r}
benchmark <- lm(lifeExp ~ gdpPercap, data = world_data)
summary(benchmark)
```

## Interpreting and Performance our Benchmark Model 
* What dose a $1000 increase in GDP Per Capita have on life expectancy? 

* How "good" is our model? What tools did you just use? 

* What about missing variables that could explain the variation in Life Expectancey?

* Our data has its limits, but econometric tool can we use to imporve the estimates? 

## Incorporating Time Fixed Effects 

```{r}
time_fe_reg <- lm(lifeExp ~ gdpPercap + factor(year), 
                  data = world_data)
summary(time_fe_reg)
```

## Did Time Fixed Effects Help?

* How do we know? What tools did you use to say yes or no? 

* Still missing variables that explain the variation between countries 

* Let's add country fixed effects!

## In Class Exercise 2: Country Fixed Effects 

* Run a regression named "country_fe_reg" with just fixed effects

* Run a regression named "both_fe_reg" with both time and country fixed effects 

* Put benchmark, time_fe_reg, country_fe_reg, and both_fe_reg into a stargazer table

* Please exclude the coefficents reported on the fixed effects, but indicate what regression has which fixed effects. 

## In Class Exercise 2: Solution

```{r, results = FALSE}
benchmark <- lm(lifeExp ~ gdpPercap, 
                data = world_data)
time_fe <- lm(lifeExp ~ gdpPercap + factor(year), 
              data = world_data)
country_fe <- lm(lifeExp ~ gdpPercap + factor(country), 
                 data = world_data)
both_fe <- lm(lifeExp ~ gdpPercap + factor(year) + 
                factor(country), data = world_data)

stargazer(benchmark, time_fe, country_fe, both_fe,
          type = "text",
          covariate.labels = c("GDP Per Capita"),
          omit = c("factor"), 
          add.lines = list(c("Time Fixed Effects?", 
                             "No", "Yes", "No", "Yes"),
                           c("Country Fixed Effects?", 
                             "No", "No", "Yes", "Yes"))
          )
```

## In Class Exercise 2: Interpreting Results

* What is the "best" model?

* Why does the sign infront of the coefficent "flip"?

* Is the model with time and country fixed effects over-fitted?

## Vizualizing Data with Maps 

* Powerful tool

* Dipicts clustering 

* We will be using another precanned data set from the 'maps' package

## Loading in Map Data 

```{r}
world <- map_data("world")
head(world)
```


## Geographic Data in R 

* Think of the data as a bunch of points where R is smart enough to just draw lines through the points

* Ordering matters in this type of geospatial data - so don't go too crazy on it!!!

* There are many types of way to store geographic data, and the type of data we are working with is the easiest.

* Just be careful if you are wanting to do maps in the future, most of the time you will be given shape files which are its own special thing.

## Geographic Data in R pt. 2

* Mapping, in this lecture, works the exact same as a normal ggplot

* There is a new "layer" called polygon

* Note that the x variable is longitude and the y variable is lattitude

* It common for people to say "latt, long" instead of "long, latt", either way to say it is fine, but when working with geographic data 90% of the time your x variable will be long and your y variable will be latt

## Our First Map

```{r}
world %>% 
    ggplot(aes(x = long, y = lat, group = group)) +
    geom_polygon(fill = "gray", color = "black", 
                 size = 0.3)
```

## Our First Map - Improvements 

* What are some improvements we can make to this map?

* Since we are using ggplots we can make the same improvements in the same manner 

* The only tricky part is removing the axes 

## Our First Map - Removing the Axes 

```{r, results="hide"}
no_axes <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank())
world %>% 
    ggplot(aes(x = long, y = lat, group = group)) +
    geom_polygon(fill = "gray", color = "black", 
                 size = 0.3) +
    no_axes

```

## Our First Map - Pretty! 

```{r, echo = FALSE}
no_axes <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank())

world %>% 
    ggplot(aes(x = long, y = lat, group = group)) +
    geom_polygon(fill = "gray",
                 color = "black", size = 0.3) +
    no_axes

```


## Merging Our Data with the World Map DF

* We will merge by country! 

* For now let's just look at Africa 

```{r}
world_data$country <- as.character(world_data$country)

africa <- world_data %>%
  filter(continent == "Africa") %>%
  inner_join(world, by = c("country" = "region"))
```


## Plotting our Merged Data 

```{r}
africa %>% 
  filter(year == 2007) %>%
  ggplot() + 
    geom_polygon(aes(long, lat, group = group), fill = "grey", color = "black") +
    no_axes
```

## Why are there missing countries? 

* Any ideas? 

* Did some get dropped in the merge? 

* Our data on GDP Per Capita and Life Expectancy goes back to 1954...

* This is fine for now! 

## Let's add some data! 

* Now let's create a heat map of life expectancy 

```{r,  results="hide"}
africa %>% 
  filter(year == 2007) %>%
  ggplot() + 
    geom_polygon(aes(long, lat, group = group, fill = lifeExp), color = "black") +
    no_axes
```


## Heat map of Life Expectancy 

```{r,  echo = FALSE}
africa %>% 
  filter(year == 2007) %>%
  ggplot() + 
    geom_polygon(aes(long, lat, group = group, fill = lifeExp), color = "black") +
    no_axes
```

## Improvements? 

* How can we make the map better? 

## Improved Heat Map 

```{r,  echo = FALSE}
africa %>% 
  filter(year == 2007) %>%
  ggplot() + 
    geom_polygon(aes(long, lat, group = group, fill = lifeExp), color = "black") +
    no_axes +
    ggtitle("Heat Map of Life Expectancy in Africa") +
    theme(plot.title = element_text(hjust = 0.5)) +
    guides(fill = guide_legend(title = "Life Expectancy")) 
    #scale_fill_gradient(colours = jet.colors) #changing the color scale is being a bit weird 
```

## In Class Exercise 3: European Heat Map 

* Please plot a heat map of the European's Countries Per Capita GDP for the year 2007


## In Class Exercise 3: Code Solution 

```{r, results = "hide"}
world_data %>%
  filter(continent == "Europe") %>%
  inner_join(world, by = c("country" = "region")) %>%
  filter(year == 2007) %>%
  ggplot() + 
    geom_polygon(aes(long, lat, group = group, fill = gdpPercap), color = "black") +
    no_axes +
    ggtitle("Heat Map of GDP PC for Europe") +
    theme(plot.title = element_text(hjust = 0.5)) +
    guides(fill = guide_legend(title = "GDP Per Capita"))
#Need to fix GB, variable type, and color scale
```
## In Class Exercise 3: Code Solution 

```{r, echo = FALSE}
world_data %>%
  filter(continent == "Europe") %>%
  inner_join(world, by = c("country" = "region")) %>%
  filter(year == 2007) %>%
  ggplot() + 
    geom_polygon(aes(long, lat, group = group, fill = gdpPercap), color = "black") +
    no_axes +
    ggtitle("Heat Map of GDP PC for Europe") +
    theme(plot.title = element_text(hjust = 0.5)) +
    guides(fill = guide_legend(title = "GDP Per Capita"))
#Need to fix GB, variable type, and color scale
```

