---
title: "Module 1"
output: beamer_presentation 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Day 2

## Recap last week

* `read_csv()`
* `ggplot()`, `aes()`, `geom_line()`, `geom_point()`, `geom_density()`
* `filter()`, `select()`, `mutate()`

## Recap Exercise 1

* Set your working directory
* Read in the data file "acs_2016_age.csv"
* Create a new data frame filtered to ages between 25-55
* Make a line plot with AGE on the x axis and INCTOT on the y axis
* Label your chart appropriately 

```{r, echo = FALSE, message=FALSE}
acs_2016_age <- read_csv("acs_2016_age.csv")
```

## Recap last week

```{r, echo = FALSE, fig.width=6, fig.height=4}
ggplot(filter(acs_2016_age, AGE <= 55, AGE >= 25)) +
  geom_line(aes(x = AGE, y = INCTOT)) +
  labs(x = "Age",
       y = "Total Income",
       title = "Average Total Income by Age (25-55)",
       caption = "Source: 2016 ACS")
```

## Recap Exercise 2

* Add a variable to the original data frame that is the gap between total and wage income
* Make a line plot with AGE on the x axis and inc_gap on the y axis
* Label your chart appropriately 


## Recap last week

```{r, echo = FALSE, fig.width=6, fig.height=4}
ggplot(acs_2016_age) +
  geom_line(aes(x = AGE, y = INCTOT - INCWAGE)) +
  labs(x = "Age",
       y = "Total - Wage Income",
       title = "Average Difference between Total and Wage Income by Age",
       caption = "Source: 2016 ACS")
```

## Goals for Today

Economics:

* Discover a relationship between wages and education

## Goals for Today

R:

* Renaming variables 
* Recoding categorical variables
* The %>% operator
* Sorting data
* Sequences
* group_by() and summarise()

## rename()

* Start by reading in "acs_2016_sample.csv"

```{r, message=FALSE}
acs_2016_sample <- read_csv("acs_2016_sample.csv")
```

* Using the glimpse(), head(), or str() function look at the variable names. 
* Are they intuitive? Easy to understand?

## Exercise: rename()

```{r, eval = FALSE}
df <- rename(df, new_variable_name = old_variable_name)
```

Create a new data frame called acs_2016_cleaned and change the names:

* STATEFIP = state_code
* SEX = sex
* AGE = age
* RACE = race
* HISPAN = hispanic
* EDUC = education
* EMPSTAT = employment_status
* INCTOT = total_income
* INCWAGE = wage_income
* UHRSWORK = hrs_worked
* WKSWORK2 = weeks_worked

Now that we've renamed the variables, let's clean up the data frame!

```{r, echo = FALSE}
acs_2016_cleaned <-
  rename(acs_2016_sample,
         state_code = STATEFIP,
         sex = SEX,
         age = AGE,
         race = RACE,
         hispanic = HISPAN,
         education = EDUC,
         employment_status = EMPSTAT,
         total_income = INCTOT,
         wage_income = INCWAGE,
         hrs_worked = UHRSWORK,
         weeks_worked = WKSWORK2)
```


## Exercise: select()

From acs_2016_cleaned select:

* state_code
* sex
* age
* race
* hispanic
* education
* employment_status
* total_income
* wage_income
* hrs_worked
* weeks_worked

```{r, echo = FALSE}
acs_2016_cleaned <-
  select(acs_2016_cleaned,
         state_code, sex, age, race, hispanic, education, 
         employment_status, total_income, wage_income, 
         hrs_worked, weeks_worked)
```


## Exercise: filter()

Filter acs_2016)cleaned to observations where:

* total_income <= 1,000,000
* wage_income <= 1,000,000
* age >= 25 & age <= 55
* hrs_worked > 0

```{r,echo = FALSE}
acs_2016_cleaned <-
  filter(acs_2016_cleaned,
         total_income <= 1000000,
         wage_income <= 1000000,
         age >= 25 & age <= 55,
         hrs_worked > 0)
```

We can begin to get an understanding of the variables in our data frame using the summary() function.

## Exercise: mutate(), ifelse()

Using mutate() and ifelse() recode the sex variable so that 

* 1 -> "Male"
* 2 -> "Female"

```{r,echo = FALSE}
acs_2016_cleaned <-
  mutate(acs_2016_cleaned,
         sex = ifelse(sex == 1, "Male", "Female"))
```

## Exercise: Mean Income by Sex

Now that we've cleaned our data frame, we can run some summary statistics.

A simple question: What is the average income for men?

* Create a new data frame that only contains observations for men called male_data
* select() the column total_income
* find the mean() and median()

```{r, echo=F}
male_data <- filter(acs_2016_cleaned, sex == "Male") 
male_data <- select(male_data, total_income)
```

## Mean Income by Sex

```{r, echo=T}
mean(male_data$total_income, na.rm = T)
median(male_data$total_income, na.rm = T)
```

* What does it mean that our average income and our median income are so different? 
* Why is the average income so much higher?

## %>%

To write more efficient code we can use nested functions f(g(x))

```{r}
nested_data <- select(filter(acs_2016_cleaned, 
                             sex == "Male"), total_income)

mean(nested_data$total_income, na.rm = TRUE)
```

We get the same answer, but this is still not great. 

## %>%

Luckily we have another solution, the pipe operator: `%>%`

```{r}
pipe_data <- 
  acs_2016_cleaned %>% 
  filter(sex == "Male") %>% 
  select(total_income)

mean(nested_data$total_income, na.rm = T)
```

## %>%

The pipe operator says: "Take the left side of the pipe and make it the first argument on the right side"

```{r, eval=F, echo=T}
filter(acs_2016_cleaned, sex == "Male")
```

is the same as

```{r, eval=F, echo=T}
acs_2016_cleaned %>%
  filter(sex == "Male")
```

* By default the %>% will put the left side as the first argument
* We can control this with the dot `(.)` as a placeholder

## %>%

Let's use seq() as an example:
```{r, eval = FALSE}
seq(start, end, interval)
```

```{r}
seq(25, 30, 1)
25 %>% seq(30, 1) # 25 is read as the first argument
30 %>% seq(25, ., 1) # 30 is directed by the placeholder
```

## Exercise: %>%

Create a new data frame called female_data and use the %>% operator calculate the mean income for women

```{r, echo=FALSE}
female_data <-
  acs_2016_cleaned %>% 
  filter(sex=="Female") %>% 
  select(total_income)
```

```{r, echo = T}
mean(female_data$total_income, na.rm = T)
```

## summarise()
 
 * We want to be able to calculate summary statistics for groups of data (i.e. men/women, white/non-white)
 * What are the arguments to the summarise() function?

```{r, eval = FALSE}
summary_df <- summarise(df, new_variable_name = mean(variable))
```

## summarise()

* Here we are using the summarise() function to calculate the mean and median income for the male_data data frame. 

```{r}
male_data %>%
  summarise(mean_inc = mean(total_income, na.rm = T),
            median_inc = median(total_income, na.rm = T))
```


## Exercise: summarise()

* Now do the same calculation but for the female_data data frame.

```{r, echo = FALSE}
acs_2016_cleaned %>% 
  filter(sex == "Female") %>% 
  select(total_income) %>%
  summarise(mean_inc = mean(total_income, na.rm = T),
            median_inc = median(total_income, na.rm = T))
```


## group_by(), summarise()

* This was a very cumbersome process for calculating the mean and median income for two subgroups of data.
* Lucky for us dplyr knows how to understand groups!
* We can tell dplyr what our group variables are by piping group_by() to summarise()

```{r, echo=T, eval=F}
group_by(variable_name)
```

* What grouping variable should we use to find median and mean income for males *and* females?

## group_by(), summarise()

```{r}
acs_2016_cleaned %>%
  group_by(sex) %>%
  summarise(mean_inc = mean(total_income, na.rm = T),
            median_inc = median(total_income, na.rm = T))
```


## mutate(), ifelse(), case_when()

Now let's go back and clean up our data for race and education so that we can calculate some more interesting summary stats.

* Use the codebook to see what the numbers mean for Educational attainment [general version]
* How do we want to break up the education variable into categories?


## mutate(), ifelse(), case_when()

I propose:

* education <= 5 ~ "No HS Diploma"
* education == 6 ~ "HS Diploma"
* education %in% c(7,8,9) ~ "Some College"
* education == 10 ~ "College Degree"
* education == 11 ~ "Graduate Degree"



## mutate(), ifelse(), case_when()

Using ifelse statements in this case can get messy very quickly! Better to use case_when() because it does not require an else statement:

```{r}
acs_2016_cleaned <- 
  acs_2016_cleaned %>%
  mutate(education = 
           case_when(education <= 5 ~ "No HS Diploma",
                    education == 6 ~ "HS Diploma",
                    education %in% c(7,8,9) ~ "Some College",
                    education == 10 ~ "College Degree",
                    education == 11 ~ "Graduate Degree"))
```

## Exercise: mutate(), ifelse(), case_when()

* Let's use the same function to recode the race variable!
* The codebook has the definitions for Hispanic origin [general version] and Race [general version]

I propose the following categories:

* race == 1 & hispanic == 0 ~ "Non-Hispanic White"
* race == 1 & hispanic %in% c(1,2,3,4) ~ "Hispanic White"
* race == 2 ~ "Black"
* race == 3 ~ "Native American"
* race %in% c(4,5,6) ~ "Asian"
* race %in% c(7,8,9) ~ "Other/Multiracial"

Code up these categories using case_when()

```{r, echo = FALSE}
acs_2016_cleaned <- 
  acs_2016_cleaned %>%
  mutate(race = case_when(race == 1 & hispanic == 0 ~ "Non-hispanic White",
                          race == 1 & (hispanic %in% c(1,2,3,4)) ~ "Hispanic White",
                          race == 2 ~ "Black",
                          race == 3 ~ "Native American",
                          race %in% c(4,5,6) ~ "Asian",
                          race %in% c(7,8,9) ~ "Other/Multiracial"))
```

## Exercise: group_by(), summarise() 1

Now that we have our data all cleaned up, we can calculate some summary data and plot it:

* Create a data frame called wage_age that contains the median wage income by age and sex
* What grouping variables should you use?
* Make a line plot of the data with age on the x axis, median_wage on the y axis and sex on the color axis
* Be sure to label your chart appropriately

## group_by(), summarise()

```{r, echo = FALSE, fig.width=6, fig.height=4}
acs_2016_cleaned %>%
  group_by(age, sex) %>%
  summarise(median_wage = median(wage_income, na.rm = T)) %>%
  ggplot(aes(x = age, y = median_wage, color = sex)) +
  geom_line() +
  labs(x = "Age", 
       y = "Median Wage",
       color = "Sex",
       title = "Median Wage by Age and Sex",
       caption = "Source: 2016 ACS")
```

## Exercise: group_by(), summarise() 2

We can also do the same with education and gender

* Create a data frame called wage_education that contains the median wage income by education level and sex
* What grouping variables should you use?

This time we want to make a bar plot of this data. Which geom should we use?

```{r, echo = FALSE}
wage_eduction <-
  acs_2016_cleaned %>%
  group_by(education, sex) %>%
  summarise(median_income = median(wage_income, na.rm = T))
```


## geom_col()

```{r, fig.width= 6, fig.height=3}
gds_plot <- wage_eduction %>% 
    ggplot(aes(x = education, y = median_income, fill = sex)) +
    geom_col(position = "dodge")
gds_plot 
```

* What are some things we will want to change about this chart?


## scale_x_discrete()

* OUR X AXIS LABELS ARE TERRIBLE!
* We want them to go in some sort of order and not run into each other
* How can we fix this? using the `scale_x_discrete` function!
* We can change the order of the variables using the limits argument and the labels using the labels argument.

## scale_x_discrete()

```{r, fig.width= 6, fig.height=3, echo=F}
gds_plot_x <- 
  gds_plot +
  scale_x_discrete(labels = c("College Degree" = "College \nDegree",
                              "Graduate Degree" = "Graduate \nDegree",
                              "HS Diploma" = "HS Diploma",
                              "No HS Diploma" = "No \nHS Diploma",
                              "Some College" = "Some \nCollege"),
                     limits = c("No HS Diploma", "HS Diploma", "Some College",
                                "College Degree","Graduate Degree"))


gds_plot_x 
```

## scale_y_continuous()

Now I want to change the y scale to go in \$10,000 increments instead of \$25,000 increments. How?

* Using the `scale_y_continuous()` function
* The argument we want to use is breaks which we want to set to a sequence
* Please also add appropriate axis labels and a title

## scale_y_continuous()

```{r, echo=F}
gds_plot_y <- gds_plot_x +
  scale_y_continuous(breaks = seq(0, 90000, 10000)) +
  labs(x = "Age", 
       y = "Median Wage",
       color = "Sex",
       title = "Median Wage by Education and Sex",
       caption = "Source: 2016 ACS")
gds_plot_y
```


## Recap Exercise

Make the same plot but instead of using sex as the color variable use race. You will need to:

* Calculate the median wage per hour for each racial group at each education level using mutate(), group_by() and summarise()
* Assume a person works their average number of 
* Use geom_col() to create a bar chart
* Use scale_x_discrete(), scale_fill_discrete(), and scale_y_continuous() to fix up the axis
* Be sure to include labels and a title

## Recap Exercise

```{r, echo = FALSE,fig.width=6, fig.height=4}
acs_2016_cleaned %>%
  mutate(wage_per_hour = wage_income/(hrs_worked * weeks_worked)) %>%
  group_by(education, race) %>%
  summarise(median_wage = median(wage_per_hour, na.rm = T)) %>%
  mutate(race = factor(race, levels = c("Non-hispanic White","Asian", "Hispanic White","Black", 
                                  "Native American", "Other/Multiracial"))) %>%
  ggplot(aes(x = education, y = median_wage, fill = race)) +
  geom_col(position = "dodge") + 
  scale_y_continuous(breaks = seq(0, 375, 75)) +
  scale_x_discrete(labels = c("College Degree" = "College \nDegree",
                              "Graduate Degree" = "Graduate \nDegree",
                              "HS Diploma" = "HS Diploma",
                              "No HS Diploma" = "No \nHS Diploma",
                              "Some College" = "Some \nCollege"),
                   limits = c("No HS Diploma", "HS Diploma", "Some College",
                              "College Degree","Graduate Degree")) +
  labs(x = "Education", 
       y = "Median Hourly Wage",
       fill = "Race/Ethnicity",
       title = "Median Hourly Wage by Race and Education Level",
       caption = "Source: 2016 ACS")

```



