---
title: "Module 1 Day 2"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

```{r}
library(tidyverse)
```

## Recap last week

Set your working directory

```{r}
setwd(~/Module1.Day2)
```

Read in the data file "acs_2016_age.csv"

```{r, echo = FALSE}
acs_2016_age <- read.csv("acs_2016_age.csv")
```

Create a new data frame filtered to ages between 25-55

```{r}
acs_2016_age_filtered <- filter(acs_2016_age, AGE >= 25 & AGE <= 55)
```

Make a line plot with AGE on the x axis and INCTOT on the y axis. Label your chart appropriatly 

```{r}
ggplot(data = acs_2016_age_filtered) +
  geom_line(aes(x = AGE, y = INCTOT)) +
  labs(x = "Age",
       y = "Total Income",
       title = "Average Total Income by Age (25-55)",
       caption = "Source: 2016 ACS")
```

Add a variable to the original data frame that is the gap between total and wage income

```{r}
acs_2016_age_mutated <- mutate(acs_2016_age, inc_gap = INCTOT - INCWAGE)
```

Make a line plot with AGE on the x axis and inc_gap on the y axis, Label your chart appropriatly.

```{r, echo = FALSE, fig.width=6, fig.height=4}
ggplot(data = acs_2016_age_mutated) +
  geom_line(aes(x = AGE, y = inc_gap)) +
  labs(x = "Age",
       y = "Total - Wage Income",
       title = "Average Difference between Total and Wage Income by Age",
       caption = "Source: 2016 ACS")
```


## rename()

Start by reading in "acs_2016_sample.csv"

```{r}
acs_2016_sample <- read_csv("acs_2016_sample.csv")
```

Using the glimpse(), head(), or str() function look at the variable names

```{r}
glimpse(acs_2016_sample)

```

Create a new data frame called acs_2016_cleaned and change the names:

* STATEFIP -> state_code
* SEX -> sex
* AGE -> age
* RACE -> race
* HISPAN -> hispanic
* EDUC -> education
* EMPSTAT -> employment_status
* INCTOT -> total_income
* INCWAGE -> wage_income
* UHRSWORK -> hrs_worked
* WKSWORK2 -> weeks_worked

```{r, echo = FALSE}
acs_2016_cleaned <-
  rename(acs_2016_sample,
         state_code = STATEFIP,
         sex = SEX,
         age = AGE,
         race = RACE,
         hispanic = HISPAN,
         education = EDUC,
         employment_status = EMPSTAT,
         total_income = INCTOT,
         wage_income = INCWAGE,
         hrs_worked = UHRSWORK,
         weeks_worked = WKSWORK2)
```


## select()

From acs_2016_cleaned select:

* state_code
* sex
* age
* race
* hispanic
* education
* employment_status
* total_income
* wage_income
* hrs_worked
* weeks_worked

```{r, echo = FALSE}
acs_2016_cleaned <-
  select(acs_2016_cleaned,
         state_code, sex, age, race, hispanic, education, 
         employment_status, total_income, wage_income, 
         hrs_worked, weeks_worked)
```


## filter()

Filter to observations where:

* total_income <= 1,000,000
* wage_income <= 1,000,000
* age >= 25 & age <= 55
* hrs_worked > 0

```{r,echo = FALSE}
acs_2016_cleaned <-
  filter(acs_2016_cleaned,
         total_income <= 1000000,
         wage_income <= 1000000,
         age >= 25,
         age <= 55,
         hrs_worked > 0)
```

We can begin to get an understanding of the variables in out data frame using the summary() function.

```{r}
summary(acs_2016_cleaned)
```


## mutate(), ifelse()

Using mutate() and ifelse() recode the sex variable so that 

* 1 -> "Male"
* 2 -> "Female"

```{r,echo = FALSE}
acs_2016_cleaned <-
  mutate(acs_2016_cleaned,
         sex = ifelse(sex == 1, "Male", "Female"))
```

## Mean Income by Sex

A simple question: What is the average income for men?

* Create a new data frame that only contains observations for men
* select the column total_income
* find the mean() and median()

## Mean Income by Sex

```{r}
male_data <- filter(acs_2016_cleaned, sex == "Male") # only rows that are male

male_data <- select(male_data, total_income) # we are interested in a single column: Income

mean(male_data$total_income, na.rm = T) # Our summary statistic, ignore NA values
median(male_data$total_income, na.rm = T)
```

What does it mean that our average income and our median income are so different? Why is the average income so much higher?

## %>%

This is not an efficient way to write this code. To write more efficient code we can use nested functions f(g(x))

```{r}
nested_data <- select(filter(acs_2016_cleaned, sex == "Male"), total_income)

mean(nested_data$total_income, na.rm = TRUE)
```

This is still not great. Luckily we have another solution, the pipe operator: `%>%`

```{r}
pipe_data <- 
  acs_2016_cleaned %>% 
  filter(sex == "Male") %>% 
  select(total_income)

mean(nested_data$total_income, na.rm = T)
```


```{r}
seq(25, 30, 1)
25 %>% seq(30, 1) # The same as seq(25, 30, 1)
30 %>% seq(25, ., 1) # The same as seq(25, 30, 1)
```

* Using the %>% operator calculate the mean income for women

```{r, echo = FALSE}
female_data <-
  acs_2016_cleaned %>% 
  filter(sex=="Female") %>% 
  select(total_income)

mean(female_data$total_income, na.rm = T)
```

## summarize()
 
Use the summarize() function to calculate the mean and median income for the male_data dataframe. Don't forget about (na.rm = T)

```{r}
male_data %>%
  summarise(mean_inc = mean(total_income, na.rm = T),
            median_inc = median(total_income, na.rm = T))
```

Now do the same calculation but for the female_data data frame.

```{r, echo = FALSE}
female_data %>%
  summarise(mean_inc = mean(total_income, na.rm = T),
            median_inc = median(total_income, na.rm = T))
```


## group_by(), summarise()

```{r}
acs_2016_cleaned %>%
  group_by(sex) %>%
  summarise(mean_inc = mean(total_income, na.rm = T),
            median_inc = median(total_income, na.rm = T))
```

* Now let's go back and clean up our data for race and education so that we can calculate some more interesting summary stats.

## mutate(), ifelse(), case_when()

Nested ifelse statements get super messy very quickly! Better to use case_when:

```{r}
acs_2016_cleaned <- 
  acs_2016_cleaned %>%
  mutate(education = case_when(education <= 5 ~ "No HS Diploma",
                               education == 6 ~ "HS Diploma",
                               education %in% c(7,8,9) ~ "Some College",
                               education == 10 ~ "College Degree",
                               education == 11 ~ "Graduate Degree"))
```

Let's use the same function to recode the race variable:

RACE		Race [general version]

* 1		White
* 2		Black/African American/Negro
* 3		American Indian or Alaska Native
* 4		Chinese
* 5		Japanese
* 6		Other Asian or Pacific Islander
* 7		Other race, nec
* 8		Two major races
* 9		Three or more major races

HISPAN		Hispanic origin [general version]

* 0		Not Hispanic
* 1		Mexican
* 2		Puerto Rican
* 3		Cuban
* 4		Other
* 9		Not Reported

I propose the following categories:

* race == 1 & hispanic == 0 ~ "Non-Hispanic White"
* race == 1 & hispanic %in% c(1,2,3,4) ~ "Hispanic White"
* race == 2 ~ "Black"
* race == 3 ~ "Native American"
* race %in% c(4,5,6) ~ "Asian"
* race %in% c(7,8,9) ~ "Other/Multiracial"

Code up these categories using case_when

```{r, echo = FALSE}
acs_2016_cleaned <- 
  acs_2016_cleaned %>%
  mutate(race = case_when(race == 1 & hispanic == 0 ~ "Non-hispanic White",
                          race == 1 & hispanic %in% c(1,2,3,4) ~ "Hispanic White",
                          race == 2 ~ "Black",
                          race == 3 ~ "Native American",
                          race %in% c(4,5,6) ~ "Asian",
                          race %in% c(7,8,9) ~ "Other/Multiracial"))
```

## group_by(), summarise()

Now that we have our data all cleaned up that calculate some summary data and plot it:

* Create a data frame called wage_age that contains the median wage income by age and sex
* What grouping variables should you use?
* Make a line plot of the data with age on the x axis, median_wage on the y axis and sex on the color axis
* Be sure to label your chart appropriatly


```{r, echo = FALSE, fig.width=6, fig.height=4}
wage_age <-
acs_2016_cleaned %>%
  group_by(age, sex) %>%
  summarise(median_wage = median(wage_income, na.rm = T))


ggplot(wage_age, aes(x = age, y = median_wage, color = sex)) +
  geom_line() +
  labs(x = "Age", 
       y = "Median Wage",
       color = "Sex",
       title = "Median Wage by Age and Sex",
       caption = "Source: 2016 ACS")
```


Now that we have our data all cleaned up that calculate some summary data and plot it:

* Create a data frame called wage_education that contains the median wage income by education level and sex
* What grouping variables should you use?

We want to make a bar plot of this data

```{r, echo = FALSE}
wage_eduction <-
  acs_2016_cleaned %>%
  group_by(education, sex) %>%
  summarise(median_income = median(wage_income))
```


## geom_col()

```{r, fig.width= 6, fig.height=3}
gds_plot <- wage_eduction %>% 
    ggplot(aes(x = education, y = median_income, fill = sex)) +
    geom_col(position = "dodge")

gds_plot 
```

What are some things we will want to change about this chart?

## scale_x_discrete()

```{r, fig.width= 6, fig.height=3}
gds_plot_x <- 
  gds_plot +
  scale_x_discrete(labels = c("College Degree" = "College \nDegree",
                              "Graduate Degree" = "Graduate \nDegree",
                              "HS Diploma" = "HS Diploma",
                              "No HS Diploma" = "No \nHS Diploma",
                              "Some College" = "Some \nCollege"),
                     limits = c("No HS Diploma", "HS Diploma", "Some College",
                                "College Degree","Graduate Degree"))


gds_plot_x 
```

## scale_y_continous()

Now I want to change the y scale to go in \$10,000 increments instead of \$25,000 increments. How?

* Using the scale_y_continous() function
* The argument we want to use is breaks which we want to set to a sequence

```{r}
gds_plot_y <- gds_plot_x +
  scale_y_continuous(breaks = seq(0, 90000, 10000)) +
  labs(x = "Age", 
       y = "Median Wage",
       color = "Sex",
       title = "Median Wage by Education and Sex",
       caption = "Source: 2016 ACS")
```

* Please also add appropriate axis labels and a title

## Recap Exersise

Make the same plot but instead of using sex as the color variable use race. You will need to:

* calculate the median wage per hour for each racial group for each education level using mutate(), group_by() and summarise()
* Assume a person's hours is how much they work per week
* use geom_col() to create a bar chart
* use scale_x_discrete() and scale_y_continous() to fix up the axes
* be sure to include lable and a title

## Recap Exersise

```{r, echo = FALSE,fig.width=6, fig.height=4}
acs_2016_cleaned %>%
  mutate(wage_per_hour = wage_income/(hrs_worked * weeks_worked)) %>%
  group_by(education, race) %>%
  summarise(median_wage = median(wage_per_hour, na.rm = T)) %>%
  mutate(race = factor(race, levels = c("Non-hispanic White","Asian", "Hispanic White","Black", 
                                  "Native American", "Other/Multiracial"))) %>%
  ggplot(aes(x = education, y = median_wage, fill = race)) +
  geom_col(position = "dodge") + 
  scale_y_continuous(breaks = seq(0, 375, 75)) +
  scale_x_discrete(labels = c("College Degree" = "College \nDegree",
                              "Graduate Degree" = "Graduate \nDegree",
                              "HS Diploma" = "HS Diploma",
                              "No HS Diploma" = "No \nHS Diploma",
                              "Some College" = "Some \nCollege"),
                   limits = c("No HS Diploma", "HS Diploma", "Some College",
                              "College Degree","Graduate Degree")) +
  labs(x = "Education", 
       y = "Median Hourly Wage",
       fill = "Race/Ethnicity",
       title = "Median Hourly Wage by Race and Education Level",
       caption = "Source: 2016 ACS")

```


