---
title: "Module 2"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, size = "tiny", width = 50, fig.height = 3.5,
                      fig.width = 5)
library(tidyverse)
library(broom)
setwd("//mq/home/m1icd00/Class_Materials/Spring_2018/module_2/day_4/")
```

# Day 4

## Recap Last Week
* Discussed long and wide data
    + Reshaped the data using `spread()` and `gather()`
* Created our own set of functions to calculate distance to the nearest metro
    + Wrapper functions
* Discussed function safety
    + `require()`, `warning()`, `stop()` 


## Recap Last Week

```{r, echo = FALSE, warning=FALSE}
joined_data <- read_csv("joined_data_v2.csv") 

joined_data %>% 
    mutate(distance_tenth = round(metro_distance, 0.1)) %>%     
    group_by(distance_tenth) %>% 
    summarize(price = mean(PRICE/SQUARE.FEET, na.rm = T)) %>% 
ggplot(aes(x = distance_tenth, y = price)) +
    geom_line() +
    labs(title = "Price vs distance from metro station",
         y = "USD/Sq ft.",
         x = "Miles",
         caption = "Source: Redfin and WMATA")+
  theme_minimal()
```


## Goals for Today

R:

* Review liner models and dates
* Create non-linear models in R

Economics:

* Further exploring the relationship between distance from a metro stop and home prices
* Investigate non-linear relationship of location, distance, time on the market on home prices
* Familiarize ourselves with different methods to account for non-linear effects

## Brief Review of Regressions

* Use it when we want to indentify a causal relationship
* Regression analysis is used to describe the relationship between:
    + A single response variable $Y$ and
    + One or more predictor variables $X_1$, $X_2$, $X_3$, ..., $X_n$
* Response should be continuous (but doesn't have to be!)
* Predictor variables can be continous, discrete, or categorical
    + Dummy variables are used to model categorical data 

## Creating an OLS Model

* Let's load in the data and see if there is a causal relationship between distance from the metro and home prices

```{r, eval = FALSE}
library(tidyverse)

setwd( ) # Put your  file directory here
joined_data <- read_csv("joined_data_v2.csv")
```

* What function do we use to create a regression model?
* What function do we use to create our regression tables? 

## The Effect of Metro Distance on Home Prices

\scriptsize
```{r, results='asis'}
library(stargazer)

dist_OLS <- lm(PRICE/SQUARE.FEET ~ metro_distance, data = joined_data)
stargazer(dist_OLS, header = F, dep.var.caption = "",
          title = "Impact of Distance From Metro on Price",
          omit.stat = c("ser", "f"),
          no.space = T)
```

## Interpreting the Results 

* How do we interpret the coefficient for `metro_distance`?
* Does our intercept have any meaningful interpretation?
* Is our current model a good model?
    + If not, how could we improve it?
    
## Adding States to the Model

* What are factors that affect housing prices that could differ from state to state?

```{r}
state_dist <- lm(PRICE/SQUARE.FEET ~ metro_distance +
                                     STATE, 
                 data = joined_data)
```

## Comparing the Results
```{r, echo = FALSE, results='asis'}
stargazer(dist_OLS, state_dist, header = F, dep.var.caption = "",
          title = "Impact of Distance From Metro on Price",
          omit.stat = c("ser", "f"),
          no.space = T)
```

## Comparing the Results

* How do we interpret the coefficients on our dummy variables?
* What happened to our `metro_distance` coefficient when we added our dummy variables?

## Days on the Market

* Conventional wisdom says that homes that spend a long time on the market generally cost less
    + Seller may have set the price to high originally
    + Buyers may be waiting for the seller to lower the price
* Our data has the list date and the sale date
    + Work with dates to create a new variable `days_on_market`
    
## Dates Review

* In R, dates are just numbers displayed in a special format
    + Generally stored as strings in our data
    + Use `as.Date()` to convert them from strings to date objects

```{r}
date1 <- as.Date("March-26-2018", format = "%B-%d-%Y")
date2 <- as.Date("Wednesday- Mar 28, 2018", 
                 format = "%A- %b %d, %Y")
```

* We are able to read dates in virtually any format that they are entered
    + We can find a list of formats [here](https://www.statmethods.net/input/dates.html)

## Dates Review

* Since dates are just numbers we can use them with mathematical functions and operations
```{r}
date2 - date1
date1 - 3
```

* We can also generate date sequences
```{r}
seq(date1, date2, by = "day")
```

## Refresher Exercise

* Load in the dates.csv data file then:
    + Combine the three variables `Year`, `Month`, and `Day` into a single string variable `Date`
    + Convert `Date` into a date variable
    
```{r, echo = FALSE, eval = FALSE}
read_csv("dates.csv") %>%
  mutate(Date = paste(Year, Month, Day)) %>%
  mutate(Date = as.Date(Date, format = "%y %b %d"))
```

## Creating days_on_market

* Remember dates are just numbers! 
* We can calculate the number of days on the market by merely subtracting the sale date by the list date and converting it back to a number

\scriptsize
```{r}
joined_data <- joined_data %>% 
    mutate(SOLD.DATE = as.Date(SOLD.DATE, format = "%B-%d-%Y"),
           LIST.DATE = as.Date(LIST.DATE, format = "%A- %b %d, %Y"),
           days_on_market = as.numeric(SOLD.DATE - LIST.DATE))
```

## Graphing the Relationship
```{r, echo = FALSE, warning=FALSE}
joined_data %>%
  mutate(days_three = round(days_on_market/3)*3) %>% 
  group_by(days_three) %>% 
  summarize(price = mean(PRICE/SQUARE.FEET, na.rm = T)) %>% 
  ggplot(aes(x = days_three, y = price)) +
    geom_line() +
    labs(title = "Price vs Days on the Market",
         y = "Average USD/Sq ft.",
         x = "Days on the Market",
         caption = "Source: Redfin and WMATA")+
    theme_minimal()
```

## Testing Our Hypothesis

* The relationship doesn't appear to be as straightforward as we initially thought
* Create a regression model that investigates the relationship between the days a house spends on the market and the price per square foot of a home
    + You are free to use any variable in our data set
    + Use `stargazer()` to display your results
    + Why do you believe your model is a good model?
    
## Introduction to Non-Linear Models

* Made some improvements to our model, but we can still do better
* Let's think more about metro distance
    + As we move farther away from the metro, distance probably matters less
    + The effect of distance likely varies from state to state
* We are unable to investigate either of the above relationships with a simple linear model

## Looking Back at Distance
```{r, echo = FALSE, warning=FALSE}
joined_data %>% 
    mutate(price = PRICE/SQUARE.FEET) %>%     
ggplot() +
    geom_point(aes(x = metro_distance, y = price)) +
    labs(title = "Price vs Distance from Metro Station",
         y = "USD/Sq ft.",
         x = "Miles",
         caption = "Source: Redfin and WMATA")+
  scale_y_continuous(limits = c(-50, 1500)) +
  theme_minimal()
```

## Adding A Quadratic Term

* In our previous models, moving from 0.5 to 0.6 miles away from the metro has the same effect on price as 2.9 to 3.0 miles away
* Can imagine that people pay extra to live in walking distance
    + The farther we move away from the metro, the less distance affects price
* We are able to apply mathematical transformations to the continous variables in our model
    + We would like to use a quadratic term in our model

## Adding A Quadratic Term

* There are two ways we can include a quadractic term in our data:
    1. Using formula functions provided by R
    2. Constructing a new variable called `dist_sq` and adding it to the model

* First, let's try to use the caret to take the exponenet
```{r}
dist_sq <- lm(PRICE/SQUARE.FEET ~ metro_distance + 
                                  metro_distance^2,
                data = joined_data)
```

## Did It Work?

```{r, echo = FALSE, results='asis'}
stargazer(dist_OLS, dist_sq, header = F, dep.var.caption = "",
          title = "Impact of Distance From Metro on Price",
          omit.stat = c("ser", "f"),
          no.space = T)
```

## Correcting Our Formula

* When we are working with formulas in R, our math operators have different meanings
    + ^ and * are used to create interactions
* Use `I()` to apply our math operators

```{r}
dist_sq <- lm(PRICE/SQUARE.FEET ~ metro_distance + 
                                I(metro_distance^2),
                data = joined_data)
```

## Did It Work?
```{r, echo = FALSE, results='asis'}
stargazer(dist_OLS, dist_sq, header = F, dep.var.caption = "",
          title = "Impact of Distance From Metro on Price",
          omit.stat = c("ser", "f"),
          no.space = T)
```

## Manually Constructing the Variable

* We should be right but let's double check to be sure
* Create a new variable `metro_sq` and use it to create a non-linear regression model
    + Compare this new model with the previous model in `stargazer()`
    + Are the coefficients the same?

## Interpreting Our Results

* Let's look back at our regression table, do the magnitidues in our table make sense?
* Direct interpretation of our coefficients becomes much more difficult
    + The effect changes as we move farther away from a metro station
    + Should check when the effect of metro distance becomes 0
    + Can evaluate the effect at the mean metro distance
* Helpful to create a graph to visualize this relationship

## broom()
* Recall our 3 main functions from the broom package
    + `tidy()` - for creating a data frame of component statistics
    + `augment()` - for observation level statistics (like fitted values and residuals)
    + `glance()`- for model level statistics (like R-squared etc.)
* We'll want to use `augment()` to plot our fitted values


## Visulaizing the Relationship
```{r, warning=FALSE, echo = FALSE}
library(broom)
dist_sq_augmented <- augment(dist_sq, joined_data)

dist_sq_augmented %>% 
  mutate(price = PRICE/SQUARE.FEET) %>%     
ggplot() +
  geom_point(aes(x = metro_distance, y = price))+
  geom_line(aes(x = metro_distance, y = .fitted), color = "red") +
  labs(title = "Price vs Distance from Metro Station",
       y = "USD/Sq ft.", x = "Miles",
       caption = "Source: Redfin and WMATA")+
  scale_y_continuous(limits = c(-50, 1500)) +
  theme_minimal()  
```

## Visulaizing the Relationship
* Fill in the code to create the graph from the previous slide

\scriptsize
```{r, warning=FALSE, eval = FALSE}
library(broom)
dist_sq_augmented <- augment( , )

dist_sq_augmented %>% 
  mutate(price = PRICE/SQUARE.FEET) %>%     
ggplot() +
  geom_point(aes(x = , y = ))+
  geom_line(aes(x = , y = ), color = ) +
  labs(title = "Price vs Distance from Metro Station",
       y = "USD/Sq ft.", x = "Miles",
       caption = "Source: Redfin and WMATA")+
  scale_y_continuous(limits = c(-50, 1500)) +
  theme_minimal()  
```
