---
title: "module3_day2_student"
output: pdf_document
---


```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, size = "tiny", width = 50, fig.height = 3.5, fig.width = 5)

library(tidyverse)
library(plotly)
library(stargazer)
library(ggthemes)
```

## Game Plan for the Day 

* Investigate some development data 
    * Measuring growth vs. measuring development
    * Precanned objects in functions 
    * Relationship between wealth and development 

* Plotting maps in R
    * Create maps with ggplot
    * Incorporate data with maps 
  
## Packages with Precaned Data Frames 

  * Some packages include sample data frames to play around with when you install it
  
  * Examples include Starwars Movie characters, stock prices, and movie ratings 
  
  * Precanned data frames allow you skip loading data in, when you want to investigate the usage of the packages! 

## Loading in Precanned Data 
```{r}
library(gapminder)
world_data <- gapminder
head(world_data)
```

## Quick Glance at the data
```{r}
library(gapminder)
world_data <- gapminder
world_data$country <- as.character(world_data$country)
head(world_data)
```

## Quick questions about the data 

* What frequency is data reported at?

* What variable is measuring quality of life (i.e Development)?


## Development vs. Growth 

* \textbf{Economic Growth} is the study of how economies grow through technology, research, and investments innovations.

* What recent technological innovations do you think that have attributed to growth in the American economy?

* \textbf{Development Economics} is the study of how nations improve the economic, political, and social well-being of its people. 

* How can we measure development?

## The Relationship between Economic Growth and Development 

* Is there a correlation between a country's wealth and how "developed" it is?  

* How much of a country's development is due to the country's wealth?

* This is our research question for the day! 

* How would we investigate this relationship with our data? 

## Plotting the relationship between GDP Per Capita and Life Expectancy

```{r}
world_data %>%
  ggplot(aes(gdpPercap, lifeExp)) +
  geom_point() 
```

## In Class Exercise 1: Make a pretty plot!!!

* Y'all are now plotting pros! 

* Fix up the previous plot and differentiate the observations by Continent. 

* I would suggest to use scale_x_log10(), if you do not know what this function does please google it or type ?scale_x_log10() into you console 

## In Class Exercise 1: Solution

```{r}
world_data %>%
  ggplot(aes(gdpPercap, lifeExp, color = )) +
  geom_point() +
  scale_x_log10() +

  
```

## Creating a Benchmark Model 

```{r}
benchmark <- lm(lifeExp ~ log(gdpPercap), data = world_data)
stargazer(benchmark, type = 'text', header = FALSE)
```

## Interpreting and Performance our Benchmark Model 
* What dose a 1% increase in GDP Per Capita have on life expectancy? Remember that this is a level-log regression! 

* Is our coefficient statistically significant? Is it economically significant?

* What about missing variables that could explain the variation in Life Expectancy?

* Our data has its limits, but econometric tool can we use to improve the estimates? 

## Incorporating Time Fixed Effects 

```{r}
time_fe_reg <- lm(lifeExp ~ log(gdpPercap) + factor(year), 
                  data = world_data)
stargazer(time_fe_reg, header = FALSE, type = 'text')
```


## Presenting Factor Models 

```{r}
stargazer(benchmark, time_fe_reg, header = FALSE,
          type = "text",
          covariate.labels = c("Log GDP Per Capita"),
          omit = c("factor"), 
          add.lines = list(c("Time Fixed Effects?", 
                             "No", "Yes"))
          )
```

## Did Time Fixed Effects Help?

* How do we know? What tools did you use to say yes or no? 

* Still missing variables that explain the variation between countries 

* Let's add country fixed effects!

## In Class Exercise 2: Country Fixed Effects 

* Run a regression named "country_fe_reg" with just country fixed effects

* Run a regression named "both_fe_reg" with both time and country fixed effects 

* Put benchmark, time_fe_reg, country_fe_reg, and both_fe_reg into a stargazer table

* Please exclude the coefficients reported on the fixed effects, but indicate what regression has which fixed effects. 

## In Class Exercise 2: Solution

```{r}
benchmark <- lm(lifeExp ~ log(gdpPercap), 
                data = world_data)
time_fe <- lm(lifeExp ~ log(gdpPercap), 
              data = world_data)
country_fe <- lm(lifeExp ~ log(gdpPercap), 
                 data = world_data)
both_fe <- lm(lifeExp ~ log(gdpPercap), 
              data = world_data)

stargazer(benchmark, time_fe, country_fe, both_fe, header = FALSE,
          type = "text",
          covariate.labels = c("Log of GDP Per Capita"),
          omit = c("factor"), 
          add.lines = list(c("Time Fixed Effects?", 
                             "No", "Yes", "No", "Yes"),
                           c("Country Fixed Effects?", 
                             "No", "No", "Yes", "Yes"))
          )
```

## In Class Exercise 2: Interpreting Results

* What is the "best" model?

* Is the model with time and country fixed effects over-fitted?

## How do I pick what fixed effects to use?

* We could have done country fixed effects, but why did I suggest to do country fixed effects instead?

* Are some fixed effects "better" than others? 

* There is no "rule" on what type of fixed effects to use when, but there are "guide lines"...

## Vizualizing Data with Maps 

* Powerful tool

* Depicts clustering 

* We will be using another precanned data set from the 'maps' package

## Loading in Map Data 

```{r}
library(maps)

world_map <- map_data("world")
head(world_map)

```

## Geographic Data in R 

* Think of the data as a bunch of points where R is smart enough to just draw lines through the points

* Ordering matters in this type of geospatial data - so don't go too crazy on it!!!

* There are many types of way to store geographic data, and the type of data we are working with is the easiest.

## Geographic Data in R pt. 2

* Mapping, in this lecture, works the exact same as a normal ggplot

* There is a new "layer" called polygon

* Note that the x variable is longitude and the y variable is latitude

* It common for people to say "lat, long" instead of "long, lat", either way to say it is fine, but when working with geographic data 90% of the time your x variable will be long and your y variable will be lat

## Our First Map

```{r}
world_map %>% 
    ggplot(aes(x = long, y = lat, group = group)) +
    geom_polygon(fill = "gray", color = "black", 
                 size = 0.3)

```

## Our First Map - Improvements 

* What are some improvements we can make to this map?

* Since we are using ggplots we can make the same improvements in the same manner 

* The only tricky part is removing the axes 

## Our First Map - Removing the axes 

```{r}
no_axes <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank())

world_map %>% 
    ggplot(aes(x = long, y = lat, group = group)) +
    geom_polygon(fill = "gray",
                 color = "black", size = 0.3) +
    no_axes +
    coord_equal()

```


## Plotting a single continent 

* Take a look at our world_map data

* Can is there a continent field? 

* How can we fix this? 

* Luckily you know R and if you google enough there is almost always a package that does what you want!

## 'countrycode' package

* This package lets you take country names and map them to a country code, a continent, or a region 

* we will be using the countrycode() function from this package

* How do we see how a function works?

## Adding continent names with countrycode()

```{r}
# install.packages('countrycode')
library(countrycode)

world_map <- world_map %>%
  mutate(continent = countrycode(sourcevar = world_map[, "region"],
                        origin ="country.name",destination="continent"))
head(world_map)
```


## Plotting a single continent - Africa! 

```{r, echo = FALSE}
world_map %>%
  filter(continent == "Africa") %>%
  ggplot() + geom_polygon(aes(long, lat, group = group), 
                          fill = "grey", color = "black") +
  no_axes +
  coord_equal() 

```


## In Class Exercise 3: Plot your favorite continent! 

* First provide me a vector with the unique continents that we can map. How does it look?

* Provide me a dataframe of countries who have NA as their continent. Why do you think they are NA?

* Please plot your favorite continent, and change the color of the countries to your favorite color! 

## In Class Exercise 3: Solution unique ! 

```{r}

#write the code to find the unique continents that show up in the world_map data frame

```

## In Class Exercise 3: Soultion pt. 2! 

```{r}
#find countries where their continent is NA

NA_continent <- filter()
head(NA_continent)
```

## In Class Exercise 3: Soultion pt. 3! 

```{r}
world_map %>%
  filter(continent == "") %>%
  ggplot() + geom_polygon(aes(long, lat, group = group), 
                          fill = "", color = "black")   +
  no_axes +
  coord_equal() 

```

## Quick Spatial Data Style Aside 

* If you have noticed, throughout class there are certain "styles" to follow when programming in R and visualizing data 

* In this class we try to teach and follow the style laid out by Hadley Wickman (the R guru)

* Just like R, Spatial Data Science has some style rules

* One important rule, is that you should not make the color of countries or land blue! 

* Blue is reserved for water! 

## Let's add data to our maps! 

* Now that we are pros at plotting basic maps, lets add some depth to them! 

* The goal is now to visualize the GDP per capital and Life Expectancy between different countries using a \textbf{heat map}

* This means we must merge our world_map with our world_data 

* What variable will be merging on?

## Right Join or Inner Join? That is the question! 

* Assuming that we will merge the data frames with world_data being the left and world_map being the right data frame, should we use an inner or right join?

* Well lets try merging both ways and map Africa!

* From there we can pick which join we should use


## Inner Join 
```{r}
world_inner <- world_data %>%
  filter(year == 2007) %>%
  inner_join(world_map, world_data, by = c("country" = "region" ))
  
  
world_inner %>%  
  filter(continent.y == "Africa") %>%
  ggplot() + 
  geom_polygon(aes(long, lat, group = group), 
               fill = "grey", color = "black") +
  no_axes +
  coord_equal()

  
```

## Right Join 
```{r}
world_right <- world_data %>%
  filter(year == 2007) %>%
  right_join(world_map, world_data, by = c("country" = "region" ))
  
world_right %>%
  filter(continent.y == "Africa") %>%
  ggplot() + 
  geom_polygon(aes(long, lat, group = group), 
               fill = "grey", color = "black") +
  no_axes +
  coord_equal() 

  
```

## Right Join Wins - Why?

* Why are there holes in Africa when use the inner join?

* What is the difference between continent.x and continent.y? Which variable belongs to which dataframe?

* Now lets make our first heat map of life expectancy!!!

## Our first Heat map of Life Expectancy! 

```{r}
world_right %>% 
  filter(continent.y == 'Africa') %>%
  ggplot() + 
  geom_polygon(aes(long, lat, group = group, fill = lifeExp), 
               color = "black") +
  no_axes +
  coord_equal()

```

## Improvements? 

* Why are there grey parts? 

* How can we make the map better? 

* Luckily, since we are using ggplot we can make improvements in the same manner that we have been doing!

## Improved Heat Map 

```{r}
world_right %>%
  filter(continent.y == "Africa") %>%
  ggplot() + 
    geom_polygon(aes(long, lat, group = group, fill = lifeExp), color = "black") +
    no_axes +
    coord_equal() +
    ggtitle("Heat Map of Life Expectancy in Africa") +
    theme(plot.title = element_text(hjust = 0.5)) +
    scale_fill_distiller(palette = "Spectral")+
    guides(fill = guide_legend(title = "Life Expectancy"))

```

## In Class Exercise 4: European Heat Map 

* Please plot a heat map of the European countries' Per Capita GDP for the year 2007

## In Class Exercise 4: Solution 

```{r, echo = TRUE, eval=FALSE}

world_right %>%
  filter(continent.y == "") %>%
  ggplot() + 
    geom_polygon(aes(long, lat, group = group, fill = gdpPercap), color = "black") +
    no_axes +
    coord_equal(ratio = ) +
    ggtitle("Heat Map of GDP PC for Europe") +
    theme(plot.title = element_text(hjust = 0.5)) +
    scale_fill_distiller(palette = "Spectral") +
    guides(fill = guide_legend(title = "GDP Per Capita"))

