---
title: "Module 2"
date: "March 9, 2018"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, size = "tiny",
                      fig.height = 2.5,
                      fig.width = 4.5,
                      fig.align = "center", width = 50, 
                      tidy.opts=list(width.cutoff=48),tidy=TRUE)

#install.packages("stringr")
library(stringr)
library(spatstat)
library(dplyr)
library(scales)
library(ggplot2)
library(stargazer)
library(Hmisc)
library(lubridate)
library(readr)
setwd("/if/fame/afe/usr/amonken/Howard/Class_Materials/Spring_2018/module_2")
```

# Day 2

## Recap Last Week

\centerline{\includegraphics[height=3in]{./Reference/for_loop.png}}

## Simple Exercise

* Cubing every number from 1 to 8 and print the results

```{r}

for(num in    ){
    print(     **3) 
}

```
## Simple Exercise

```{r}

for(num in seq(1,8)){
    print(num**3) 
}

```

## Recap Last Week
* Like we did last class, let's convert this vector of substrings into a single string. Make sure to use meaningful variable names in your code. (Hint: Use the paste function.)

```{r, eval = FALSE}
string_vec <- c("Economics","is","the","best","subject!")

## initialize the variable

for(item in string_vec){
  ## code to be executed
}

```

```{r "In-class answer", eval=FALSE, echo=FALSE}
string_vec <- c("Economics","is","the","best","subject!")

## initialize the variable
phrase = NULL

for(item in string_vec){
  phrase = paste(phrase, item)
}

print(phrase)
```


## Goals for Today

Economics - Question of the Day

* How does the sale price of residential property change based on various factors?

### What things will affect the property value?

## Goals for Today

* Determinants of property values
    + Size of house (square footage)
    + Size of property
    + Higher floors vs lower floors
    + Quality of local schools
    + Number of bedrooms and bathrooms
    + Neighborhood location (metro, restaurants, library)
    
## Goals for Today

Programming - R

* If/Else Statements, `str_detect`, `bindrows` to append multiple datasets
* Joining multiple data frames together using dplyr `join` functions
* Cleaning data for effective data visualization

## If Else Statements
* We have previously used `ifelse()` to make decisions about recoding our variables with `mutate`
* If/else statements are work in a similar way
* In R, a basic if else statement takes the following form:
```{r, eval=FALSE}
if(logical argument){
  ## code to be executed
} else{
  ## code to be executed
}
```

## If Else Statements

\centerline{\includegraphics[height=3in]{./Reference/if_else_statement.png}}

## Evens and Odds Example
* We will use the modulo operator (`%%`) to characterize numbers as even or odd
* `%%` returns the remainder after division of the first argument by the second argument
```{r}
# %% is the modulus operator 
# We are finding the remainder!
4 %% 1; 10 %% 4; 3 %% 2
```

```{r eval=FALSE}
testNum <- 12
if(testNum %% 2 == 0){
  print(paste(testNum, "is EVEN"))
} else{
  # If the remainder after division by 2 is not 0 
  # then it must be odd (right..?)
  print(paste(testNum, "is ODD"))
}
```

## Evens and Odds
* Let's characterize and record the numbers from 1 to 10 as even or odd
```{r}
evens <- numeric()
odds <- numeric()

for(i in seq(1, 10, by = 1)){
  if(i %% 2 == 0){ # %% is the modulus operator 
                   # --- we are finding the remainder!
    evens <- c(evens, i)
  } else{
    odds <- c(odds, i)
    
  }
}
```

## Evens and Odds
```{r}
evens
odds
```

## In-Class Exercise
**Given the following grade, use an if else statement to determine if the student passed or failed**
**(cutoff is 60, with a 60 being a pass)**
```{r}
student_grade <- 71
```
**Return a correct statement that the student passed or failed**

```{r, , echo = F, eval = F, include = F}
if(   ){
  print("")
} else {
  print("")
}
```

```{r, In-class Answer, echo = F, include = F}
if(student_grade >= 60){
  print("Student passed")
} else {
  print("Student failed")
}
```

## If else statements
* We aren't limited to choosing between two conditions
* Similar to `case_when()`, else if lets us make multiple decisions
    + provides us with even more flexibility
* Checks each condition one by one
    + check the first condition, if false then it moves on to the the next one
* **Else catches everything that does not meet the previous criteria** so be careful when coding or deciding what to include

## Example
**Create a loop that will take the square root of a positive number or give us an NA if the number is negative. Save the results in a vector you initialized outside of the loop.**

```{r echo = F, include = F}
response_list <- c(0,1,2,2,0,1,2,1,0,2,2,1,1,1,2,1,0)

female <- 0
male <- 0
other <- 0

for(person in response_list){
  if(person == 1) {
    male <- male + 1
  } else if (person == 2){
    female <- female + 1
  } else {
    other <- other + 1
  }
}

survey_genders <- data.frame(male,female,other)
survey_genders
```

## Assigning Grades - In-Class Exercise

* You've been given a list of test scores that you want to categorize into letter grades
* Use if else statements to assign letters to the numeric grades
    + A 90 - 100
    + B 80 - 89
    + C 70 - 79
    + D 60 - 69
    + F < 60

```{r, eval=FALSE}
test_scores <- c(85,55,100,67,73,92,94,99,87,89.3)

# Initialize our vector
letter_grades <- 

for(grade in test_scores){
  if(   >= 90){
    letter_grades <- paste(letter_grades,"A")
  } else if(   ) {
    letter_grades <-    (             ,"B")
  } else if
    letter_grades <- 
    else if 
    
  } else {
    letter_grades <- paste(       "F")
  }
}
letter_grades
```

## Assigning Grades

```{r, eval=TRUE, echo = FALSE}
test_scores <- c(85,55,100,67,73,92,94,99,87,89.3)
# Initialize our vector
letter_grades <- NULL
for(grade in test_scores){
  if(grade >= 90){
    letter_grades <- paste(letter_grades,"A")
  } else if(grade >= 80) {
    letter_grades <- paste(letter_grades,"B")
  } else if (grade >= 70) {
    letter_grades <- paste(letter_grades,"C")
  } else if (grade >= 60) {
    letter_grades <- paste(letter_grades,"D")
  } else {
    letter_grades <- paste(letter_grades,"F")
  }
}
letter_grades
```


## Reading in the Redfin Data

* Before we can access our datasets we need to help R find them
* We can create a vector of all the files in our data folder using the `list.files` function
* *What sorts of files are in the folder?*
* Now that we know the names of the files with the datasets, we need to decide which datasets are useful and which go together
* *What are the two major categories of files?*
* *How can we combine these files most effectively?*

```{r, results='hide'}
## create a vector of dataset names
files <- list.files("./Data/")
files

```

## str_detect()
* `str_detect()` is a function in the tidyverse package that can tell whether a string contains a certain word or phrase
    + `TRUE` if the word is in the string
    + `FALSE` if the word is *not* in the string
    
\footnotesize
```{r, results='hide'}
str_detect("property_redfin_01.csv", "property") # TRUE
str_detect("location_redfin_01.csv", "property") # FALSE

str_detect(c("property_redfin_01.csv", 
             "location_redfin_01.csv"),
           "location") # c(FALSE, TRUE)
```

## In-Class Exercise
* Using what we've learned about `str_detect`, let's create a vector that only includes location files.

```{r, include = TRUE}
files[str_detect(files, "location")]
```


## Reading in the data - In-Class Exercise
* Now we have all the tools we need to build our dataset!
* Read in the location and property data using for loops and if else statements. 
    + Within the loop, combine all of the location data into one dataset, and all of the property data into one dataset.
**Hint(check out what the `bind_rows` function does.)**

\scriptsize
```{r, eval=FALSE}
property_data <- data_frame() 
location_data <- data_frame() 

for(  ){
  
    # You only want to have property data in the first dataset
    if(str_detect(   )){ 
        data <- read_csv(   )
        property_data <- bind_rows(property_data, data)
    } 
  
  # Now let's combine location data into a different dataset
  else if(str_detect(file, "location")){ 
        data <- read_csv(file)
        location_data <- bind_rows(location_data, data)
    }
}
```

\scriptsize
```{r echo=FALSE, message=FALSE}
property_data <- data_frame() # initialize empty data frame
location_data <- data_frame() # initialize empty data frame

for(file in files){
# this is a property file, add to property_data
    if(str_detect(file, "property")){ 
        data <- read_csv(file)
        property_data <- bind_rows(property_data, data)
    } 
  # location file
  else if(str_detect(file, "location")){ 
        data <- read_csv(file)
        location_data <- bind_rows(location_data, data)
    }
}
```


## Overview of the Data
* Let's take a look at the variables in our two dataframes - property data and location data
* The good news is that we have much of the data we need, the bad news is that the data is split between location information and price information


## "Joining" the Data

* Combining data frames is called "joining." There are multiple types of joins as you can see in the dplyr cheat sheet:
* Help -> cheat sheets -> Data transformation with dplyr.

\centerline{\includegraphics[height=2.7in]{./Reference/joins.png}}

## "Joining" the Data

* Harnessing the power of the `%in%` function

\scriptsize
```{r}
vec1 <- c(1, 5, 8, 10, 3, 7, 9)
vec2 <- c(4, 5, 19, 10, 5, 1, 8)
vec1 %in% vec2
which(vec1 %in% vec2) 
# position in vec1 with overlapping elements in vec2
vec1[vec1 %in% vec2] 
# we can even get back those elements that overlap
```

## "Joining" the Data

* Now using the example we just discussed as a guide, find the overlapping variables in our property and location datasets.
* Make sure to save that overlap result into a variable so we can use it for our next step.

```{r, echo = FALSE, eval = TRUE, include = FALSE}
property_names <- names(property_data)
location_names <- names(location_data)

property_names %in% location_names

overlap <- property_names[property_names %in% location_names]
overlap


```

## "Joining" the Data

* Our URL column exists in both of the dataframes!
* We only want houses that include location data...
* Which type of joining would we need to accomplish this?


## "Joining" the Data
\centerline{\includegraphics[height=3in]{./Reference/joins.png}}

## "Joining" the Data

\footnotesize
```{r}
joined_data <- full_join(property_data,
                         location_data,
                         by = overlap)

names(joined_data)
```

## "Joining" the Data

* Success! Now all of our housing data is combined
* Note - It is also possible to perform a join where the columns we are matching in each table do not have the same name using the by argument: by = c("name1" = "name2").
* Similarly, it is also possible to join on multiple columns: by = c(column1, column2, etc...)

## Examining the Data

* Now we can examine the data to check for any problems.
* What are some problems you see in terms of usability for this data?
```{r include = F, eval = F}
View(joined_data)
```

## Examining the Data

### Lots of Issues Here

* city_state column - state is inconsistent (va, VA, Virginia) and unwanted states
* propertytype column - has "propertytype" in front of every observation
* zip_code column - extra zeros
* lat_lon column - & in the middle of the string
* sold.dates, list.date columns - not date objects

### So let's first make a scatter-plot of our home price vs square feet variables.

## Cleaning the Data

```{r echo= FALSE}
joined_data %>% 
    ggplot(aes(x = SQUARE.FEET, y = PRICE)) +
    geom_point()
```

## Cleaning the Data

\footnotesize
```{r}
head(joined_data$PRICE)
```

* There's a USD in front of each numeric part of the column
* What class of variable is the price column?

## Cleaning the Data

* How can we fix this problem for a string variable? What function could we use?

## Cleaning the Data - In-Class Exercise

* One way we could do this is by replacing the USD with nothing, "". This would have the same effect as removing it.
* Try using the str_replace function to modify the string "USD40000" to be "40000"
```{r eval = F, echo = F}
str_replace()
```

```{r eval = T, echo = F, include = FALSE}
str_replace("USD40000", "USD", "")
```

## Cleaning the Data

* Let's mutate the price column to remove "USD" from the character string

\scriptsize
```{r eval = T, echo = T, results = F}
joined_data <- joined_data %>% 
    mutate(PRICE = str_replace(PRICE, "USD", ""))
head(joined_data$PRICE)
```

## Cleaning the Data - In-Class Exercise

* We just used str_replace to fix our PRICE column. 
* Check out the PROPERTY.TYPE column and use `str_replace` to fix it

```{r "in-class exercise answer", echo = F, include = F}
joined_data <- joined_data %>% 
    mutate(PROPERTY.TYPE = str_replace(PROPERTY.TYPE,
    "propertytype:", ""))
```


## Cleaning the Data

* Let's redo the plot!
```{r, echo= FALSE, fig.width=4.5, fig.height=2.5}
joined_data %>% 
    ggplot(aes(x = SQUARE.FEET, y = PRICE)) +
    geom_point() +
    labs(title = "Relationship between price and home size",
         y = "USD",
         x = "Sq ft",
         caption = "Data from Redfin")
```

## Cleaning the Data - In-Class Exercise
* We still have the same problem, we forgot to convert our PRICE column to be class numeric!
* Convert the price column to be numeric
    + remove any missing values from the price and square feet columns.
    + Then replot and see how it looks
    
## Cleaning the Data
```{r, echo= FALSE, fig.width=4.5, fig.height=2.5}
joined_data <- joined_data %>% 
    filter(!is.na(PRICE), !is.na(SQUARE.FEET)) %>%
    mutate(PRICE = as.numeric(PRICE))

# Now make the plot
joined_data %>% 
    ggplot(aes(x = SQUARE.FEET, y = PRICE)) +
    geom_point() +
    labs(title = "Relationship between price and home size",
         y = "USD",
         x = "Sq ft",
         caption = "Data from Redfin")
```
* I want to change the scale so we don't have to deal with exponentials

## Cleaning the Data

* By manually entering our scale for the y-axis, we can fix the problem with exponentials on the y-axis
* `scale_y_continuous("USD", labels = dollar)`

```{r, message = F, echo= FALSE, eval = FALSE, fig.width=4.5, fig.height=2.5, warning = FALSE}
#install.packages("scales")
library(scales)

joined_data %>% 
    ggplot(aes(x = SQUARE.FEET, y = PRICE)) +
    geom_point() +
    scale_y_continuous("USD", labels = dollar) + 
    labs(title = "Relationship between price and home size",
         x = "Sq ft",
         caption = "Data from Redfin")
```

## Cleaning the Data
* What relationship is there between price and home size?
```{r, message = F, echo= FALSE, fig.width=4.5, fig.height=2.5, warning = FALSE}
#install.packages("scales")
library(scales)

joined_data %>% 
    ggplot(aes(x = SQUARE.FEET, y = PRICE)) +
    geom_point() +
    scale_y_continuous("USD", labels = dollar) + 
    labs(title = "Relationship between price and home size",
         x = "Sq ft",
         caption = "Data from Redfin")
```



## Understanding the Data

* What about property type and price?
* Is there a relationship between these two variables?
* Produce a table of the average price per square foot by property type and make a bar plot of it.

## Understanding the Data

```{r}
type_price <- joined_data %>% 
    group_by(PROPERTY.TYPE) %>% 
    summarise(price = mean(PRICE/SQUARE.FEET, na.rm = T))

type_price
```

## Understanding the Data
Luckily this is not a production-quality graph, just something for our reference.

* How does theme_light change the graph? How is color assigned?

```{r, echo = FALSE}
type_price %>% 
            ggplot(aes(x = PROPERTY.TYPE, 
                       y = price, 
                       fill = PROPERTY.TYPE)) +
            geom_bar(stat = "identity") +
            scale_y_continuous("USD", labels = dollar) +
            scale_fill_discrete(guide = F) + # get rid of the fill legend
            theme_light() # Change the overall aesthetic
```

## Understanding the Data

* City/state is another variable that might impact home sale price... Why?

```{r}
unique(joined_data$CITY_STATE)
```
* What issues do we have with the city_state variable?

## Cleaning the Data

* We can use the `str_split_fixed` function to alter our string (or vector of strings) using a pattern
* The function returns the peices of the string after splitting based on your pattern

```{r}
strings <- str_split_fixed("United States of America", "of", n = 2)

strings
```

## Cleaning the Data

* We told R to split into two piece (n = 2), so we got two columns in return.
* If we want to get back to individual elements, we can subset the variable.

```{r}
strings[, 1] # first row, first column
strings[, 2] # first row, second column
```

## Cleaning the Data

* Let's test it out with something closer to our actual use case.

```{r}
str_split_fixed("Arlington, Virginia", ",", n = 2)
```

## Cleaning the Data - In-Class Exercise

* Use what we just learned to mutate the city_state column and create separate city and state columns.


## Cleaning the Data
```{r, echo = FALSE}
joined_data <- joined_data %>% 
    mutate(CITY = str_split_fixed(CITY_STATE, ",", n = 2)[,1],
           STATE = str_split_fixed(CITY_STATE, ",", n = 2)[, 2])

joined_data %>% select(CITY_STATE, CITY, STATE) %>% head
```

## Cleaning the Data - In-Class Exercise

* We also need to clean the LAT_LON column by splitting it into latitude and longitude columns
* Right now our column is separated by `&` character, so let's split it using the method we just used

```{r "lat_lon answer", echo = F, include = F, eval = T}
joined_data <- joined_data %>% 
    mutate(LATITUDE = str_split_fixed(LAT_LON, "&", n = 2)[,1],
           LONGITUDE = str_split_fixed(LAT_LON, "&", n = 2)[,2])
```

## Cleaning the Data
* We have one more problem - there are spaces around our new state variable

```{r}
head(joined_data$STATE)
```

* The function `str_trim` can be used to fix this

## Cleaning our Data

* All the blank spaces are removed!
```{r}
str_trim(" word ") # It removes the blank space!
```
* Now let's do it for our data

## Cleaning the Data

* Voila! All blank spaces gone.

\scriptsize
```{r}
joined_data <- joined_data %>% 
    mutate(STATE = str_trim(STATE))

head(joined_data$STATE)
```

## Cleaning the Data

* We're not done yet. Let's take another look at the state column and see what types of responses exist

\scriptsize
```{r}
unique(joined_data$STATE)
```

* What do you think we need to do to further clean this column?

## Cleaning the Data - In-Class Exercise

* The function `str_to_upper` will convert all lower case letter to upper case ones
```{r}
str_to_upper("va")
```

* Now use this function to mutate the state column to upper case

## Cleaning the Data
\scriptsize
```{r}
joined_data <- joined_data %>% 
    mutate(STATE = str_to_upper(STATE))

unique(joined_data$STATE)
```

## Cleaning the Data - In-Class Exercise
* Now we need to change the instances of "VIRGINIA" to "VA"
* We've already used this function! 
* Go back and find the function you need to make that change

```{r "VIRGINIA ANSWER", echo = F, include = F}
joined_data <- joined_data %>% 
    mutate(STATE = str_replace(STATE, "VIRGINIA", "VA"))
```

## Cleaning the Data - In-Class Exercise

* The last thing we need to do is toss out the rows with MICHIGAN and COLORADO since we only want data along the DC Metro
* Use the `%in%` function to filter only rows with VA, DC, or MD in them.

## Cleaning the Data

```{r}
joined_data <- joined_data %>% 
    filter(STATE %in% c("VA", "DC", "MD"))

unique(joined_data$STATE)
```

## Plotting the Data
* Let's find out the average prices by state for our dataset

```{r, eval = FALSE}
state_average <- joined_data %>% group_by(STATE) %>% 
    summarise(price = mean(PRICE/SQUARE.FEET, na.rm = T))

state_average_plot <- state_average %>% 
    ggplot(aes(x = STATE, y = price, fill = STATE)) +
    geom_bar(stat = "identity", width = 0.5) + 
    labs(x = NULL, # What does this do?
         y = "USD/Sq ft",
         title = "Average price for DC Home sales",
         caption = "Data from Redfin") 

state_average_plot
```

## Plotting the Data
* I want to have a larger plot title and have it centered
* How can we do that? Would we use a data-related dimension or an aesthetic option?

```{r, echo = FALSE}
state_average <- joined_data %>% group_by(STATE) %>% 
    summarise(price = mean(PRICE/SQUARE.FEET, na.rm = T))

state_average_plot <- state_average %>% 
    ggplot(aes(x = STATE, y = price, fill = STATE)) +
    geom_bar(stat = "identity", width = 0.5) + 
    labs(x = NULL, # What does this do?
         y = "USD/Sq ft",
         title = "Average price for DC Home sales",
         caption = "Data from Redfin") 

state_average_plot
```

## Plotting the Data
* We can use the `theme()` function to alter elements of the graph.
* What other textual elements can we alter in this graph?

\scriptsize
```{r, fig.width=4, fig.height=2}
state_average_plot +
    theme(plot.title = 
            element_text(size = 14, hjust = 0.5, color = "Red"))
```

## Plotting the Data

* Since our title, and the other labels, is a character string, or text, we use element_text to control it. element_text allows us to set:

    + font family (Times New Roman, Arial, etc...) 
    + font size (10, 12, 14, etc...)  
    + font face (bold, italic) 
    + color 
    + hjust (horizontal adjustment) 
    + vertical adjustment
    + angle  
    + other aspects that impact the display of text.

## Plotting the Data

\scriptsize
```{r, , eval = FALSE, fig.width=5, fig.height=3}
state_proptype <- joined_data %>% 
    group_by(PROPERTY.TYPE, STATE) %>% 
    summarise(price = mean(PRICE/SQUARE.FEET, na.rm = T)) # empty lots

plot <- state_proptype %>% 
    ggplot(aes(x = STATE, y = price, fill = STATE)) +
    geom_bar(stat = "identity") +
    facet_wrap("PROPERTY.TYPE") +
    labs(title = "Price by property type",
         x = NULL,
         y = "USD/Sq ft.",
         caption = "Data from Redfin")

plot + theme(plot.title = element_text(face = "italic", size = 14, color = "Blue"),
          axis.text.x = element_text(face = "bold", color = "green4"),
          plot.background = element_rect(fill = "beige", color = "red"),
          panel.background = element_rect(fill = "cadetblue1", 
                                          color = "black", linetype = 3))
              
              
```

## Plotting the Data
\normalsize
```{r, echo = FALSE, fig.width=4.5, fig.height=3}
state_proptype <- joined_data %>% 
    group_by(PROPERTY.TYPE, STATE) %>% 
    summarise(price = mean(PRICE/SQUARE.FEET, na.rm = T)) # empty lots
state_proptype %>% 
    ggplot(aes(x = STATE, y = price, fill = STATE)) +
    geom_bar(stat = "identity") +
    facet_wrap("PROPERTY.TYPE") +
    labs(title = "Price by property type",
         x = NULL,
         y = "USD/Sq ft.",
         caption = "Data from Redfin") +
    theme(plot.title = element_text(face = "italic", size = 14, color = "Blue"),
          axis.text.x = element_text(face = "bold", color = "green4"),
          plot.background = element_rect(fill = "beige", color = "red"),
          panel.background = element_rect(fill = "cadetblue1", 
                                          color = "black", linetype = 3))
              
              
```

## Understanding the Data

* We used `element_rect()` to adjust rectangular elements of the plot.
* The main plot has a beige background with a red border.
* We used the panel.background option to adjust elements of the area where the data is
* What did the linetype argument do in our panel.background option?
* What other rectangles could we adjust?

```{r, echo= FALSE, eval = FALSE, fig.width=5, fig.height=3}
state_proptype <- joined_data %>% 
    group_by(PROPERTY.TYPE, STATE) %>% 
    summarise(price = mean(PRICE/SQUARE.FEET, na.rm = T)) # empty lots

state_proptype %>% 
    ggplot(aes(x = STATE, y = price, fill = STATE)) +
    geom_bar(stat = "identity") +
    facet_wrap("PROPERTY.TYPE") +
    labs(title = "Price by property type",
         x = NULL,
         y = "USD/Sq ft.",
         caption = "Data from Redfin") +
    theme(plot.title = element_text(face = "italic", size = 14, color = "Blue"),
          axis.text.x = element_text(face = "bold", color = "green4"),
          plot.background = element_rect(fill = "beige", color = "red"),
          panel.background = element_rect(fill = "cadetblue1", 
                                          color = "black", linetype = 3))
              
              
```

## Understanding the Data
* Perhaps using the price per square foot is not the best method of analysis.
* When a property is sold the surrounding land around the house will likely affect the selling price.
* Let's work on making a plot of price/lot size

## Understanding the Data
* We'll start by looking at the head of the data
```{r, echo = FALSE}
joined_data %>% select(PRICE, SQUARE.FEET, LOT.SIZE) %>% head()
```
* We have a lot of NA values. Why is that?

## Understanding the Data - In-Class Exercise

* Let's mutate the lot.size column so that if there is an `NA` then we replace it with the value of the square.feet column and otherwise keep lot.size the same

## Understanding the Data

\scriptsize
```{r, eval = FALSE}
joined_data <- joined_data %>% 
    mutate(LOT.SIZE = ifelse(is.na(LOT.SIZE), SQUARE.FEET, LOT.SIZE))

# Now we can find our average
lot_state_data <- joined_data %>% 
    group_by(PROPERTY.TYPE, STATE) %>% 
    summarise(price = mean(PRICE/LOT.SIZE, na.rm = T))

head(lot_state_data)
```

## Understanding the Data
```{r, echo = FALSE}
joined_data <- joined_data %>% 
    mutate(LOT.SIZE = ifelse(is.na(LOT.SIZE), SQUARE.FEET, LOT.SIZE))

# Now we can find our average
lot_state_data <- joined_data %>% 
    group_by(PROPERTY.TYPE, STATE) %>% 
    summarise(price = mean(PRICE/LOT.SIZE, na.rm = T))

head(lot_state_data)
```

## Understanding the Data
* The culminating exercise of the day!
* Make a plot like the one we just did but using our new price/lot.size variable with some modifications -
    + **Make your title centered and Green**
    + **Make your panel background White with a red border**

```{r "Lot size plot answer", echo = F, include = F}
lot_state_data %>% 
    ggplot(aes(x = STATE, y = price, fill = STATE)) +
    facet_wrap("PROPERTY.TYPE") +
    geom_bar(stat = "identity", width = 0.5) +
    labs(x = NULL,
         title = "Price per square foot of total property size",
         caption = "Data from Redfin",
         y = "USD/Sq ft.") +
    theme(plot.title = element_text(hjust = 0.5, color = "green"),
          panel.background = element_rect(fill = "white", color = "red"))
```

## Understanding the Data

```{r, echo = F, fig.width=4.5, fig.height=3}
lot_state_data %>% 
    ggplot(aes(x = STATE, y = price, fill = STATE)) +
    facet_wrap("PROPERTY.TYPE") +
    geom_bar(stat = "identity", width = 0.5) +
    labs(x = NULL,
         title = "Price per square foot of total property size",
         caption = "Data from Redfin",
         y = "USD/Sq ft.") +
    theme(plot.title = element_text(hjust = 0.5, color = "green"),
          panel.background = element_rect(fill = "white", color = "red"))
```