---
title: "Module 2"
date: "March 9, 2018"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, size = "tiny",
                      fig.height = 2.5,
                      fig.width = 4.5,
                      fig.align = "center", width = 50, 
                      tidy.opts=list(width.cutoff=48),tidy=TRUE)

#install.packages("stringr")
library(stringr)
library(spatstat)
library(dplyr)
library(scales)
library(ggplot2)
library(stargazer)
library(Hmisc)
library(lubridate)
library(readr)
#install.packages("scales")
library(scales)
setwd("/if/fame/afe/usr/amonken/Howard/Class_Materials/Spring_2018/module_2")
```

# Day 2

## Recap Last Week

\centerline{\includegraphics[height=3in]{./Reference/for_loop.png}}

## Simple Exercise

* Cubing every number from 1 to 8 and print the results

```{r, eval = F}

for(num in    ){
    print(     **3) 
}

```
## Simple Exercise

```{r}

for(num in seq(1,8)){
    print(num**3) 
}

```

## Recap Last Week - In-Class Exercise
* Like we did last class, let's convert this vector of substrings into a single string. Make sure to use meaningful variable names in your code. (Hint: Use the paste function.)

```{r, eval = FALSE}
string_vec <- c("Economics","is","the","best","subject!")

## initialize the variable

for(item in string_vec){
  ## code to be executed
}

```

## Recap Last Week - In-Class Exercise (Answer)
```{r "In-class answer", eval=FALSE, echo=TRUE}
string_vec <- c("Economics","is","the","best","subject!")

## initialize the variable
phrase = NULL

for(item in string_vec){
  phrase = paste(phrase, item)
}

print(phrase)
```


## Goals for Today

Economics - Question of the Day

* How does the sale price of residential property change based on various factors?

### What things will affect the property value?

## Goals for Today

* Determinants of property values
    + Size of house (square footage)
    + Size of property
    + Higher floors vs lower floors
    + Quality of local schools
    + Number of bedrooms and bathrooms
    + Neighborhood location (metro, restaurants, library)
    
## Goals for Today

Programming - R

* If/Else Statements, `str_detect`, `bindrows` to append multiple datasets
* Joining multiple data frames together using dplyr `join` functions
* Cleaning data for effective data visualization

## If/Else Statements
* We have previously used `ifelse()` to make decisions about recoding our variables with `mutate`
* If/else statements work in a similar way
* In R, a basic if else statement takes the following form:
```{r, eval=FALSE}
if(logical argument){
  ## code to be executed
} else{
  ## code to be executed
}
```

## If Else Statements

\centerline{\includegraphics[height=3in]{./Reference/if_else_statement.png}}

## Evens and Odds Example
* We will use the modulo operator (`%%`) to characterize numbers as even or odd
* `%%` returns the remainder after division of the first argument by the second argument
```{r}
# %% is the modulus operator 
# We are finding the remainder!
4 %% 1; 10 %% 4; 3 %% 2
```

```{r eval=FALSE}
testNum <- 12
if(testNum %% 2 == 0){
  print(paste(testNum, "is EVEN"))
} else{
  # If the remainder after division by 2 is not 0 
  # then it must be odd (right..?)
  print(paste(testNum, "is ODD"))
}
```

## Evens and Odds Example
* Let's characterize and record the numbers from 1 to 10 as even or odd
```{r}
evens <- numeric()
odds <- numeric()

for(i in seq(1, 10, by = 1)){
  if(i %% 2 == 0){ # %% is the modulus operator 
                   # --- we are finding the remainder!
    evens <- c(evens, i)
  } else{
    odds <- c(odds, i)
    
  }
}
```

## Evens and Odds
```{r}
evens
odds
```

## In-Class Exercise
* Given the following grade, use an if/else statement to determine if the student passed or failed
    + **(A passing grade is greater than or equal to 60)**
```{r}
student_grade <- 71
```
**Return a correct statement that the student passed or failed**

```{r, , echo = T, eval = F}
if(             ){
  print("     ")
} else {
  print("     ")
}
```

## In-Class Exercise (Answer)
```{r, In-class Answer, echo = F, include = F}
if(student_grade >= 60){
  print("Student passed")
} else {
  print("Student failed")
}
```

## If else statements
* We aren't limited to choosing between two conditions
* Similar to `case_when()`, else if lets us make multiple decisions
    + provides us with even more flexibility
* Checks each condition one by one
    + check the first condition, if false then it moves on to the the next one
* **Else catches everything that does not meet the previous criteria** so be careful when coding or deciding what to include

## Example
* Here's an example of assigning survey numerical responses to genders.

\scriptsize
```{r}
response_list <- c(0,1,2,2,0,1,2,1,0,2,2,1,1,1,2,1,0)
female <- 0; male <- 0; other <- 0

for(person in response_list){
  if(person == 1) {
    male <- male + 1
  } else if (person == 2){
    female <- female + 1
  } else {
    other <- other + 1
  }
}
survey_genders <- data.frame(male,female,other)
survey_genders
```

## Assigning Grades - In-Class Exercise

* You've been given a list of test scores that you want to categorize into letter grades
* Use if else statements to assign letters to the numeric grades
    + A 90 - 100, B 80 - 89, C 70 - 79, D 60 - 69, F < 60

\scriptsize
```{r, eval=FALSE}
test_scores <- c(85,55,100,67,73,92,94,99,87,89.3)

# Initialize our vector
letter_grades <-   

for(grade in test_scores){
  if(   >= 90){
    letter_grades <- paste(letter_grades,"A")
  } else if(   ) {
    letter_grades <-    (            ,"B")
  } else if
    letter_grades <- 
    else if 
    
  } else {
    letter_grades <- paste(       "F")
  }
}
letter_grades
```

## Assigning Grades - In-Class Exercise (Answer)

\scriptsize
```{r, eval=TRUE, echo = TRUE}
test_scores <- c(85,55,100,67,73,92,94,99,87,89.3)
# Initialize our vector
letter_grades <- NULL
for(grade in test_scores){
  if(grade >= 90){
    letter_grades <- paste(letter_grades,"A")
  } else if(grade >= 80) {
    letter_grades <- paste(letter_grades,"B")
  } else if (grade >= 70) {
    letter_grades <- paste(letter_grades,"C")
  } else if (grade >= 60) {
    letter_grades <- paste(letter_grades,"D")
  } else {
    letter_grades <- paste(letter_grades,"F")
  }
}
letter_grades
```


## Reading in the Redfin Data

* Before we can access our datasets we need to help R find them
* We can create a vector of all the files in our data folder using the `list.files` function
* **What sorts of files are in the folder?**
* Now that we know the names of the files with the datasets, we need to decide which datasets are useful and which go together
* **What are the two major categories of files?**
* **How can we combine these files most effectively?**

\scriptsize
```{r, results='hide'}
## create a vector of dataset names
files <- list.files("./Data/")
files

## to be able to access these files, we'll need to tell R to look in the folder 'Data'
## therefore, we should append 'data' onto the title using paste0
data_files <- paste0("./Data/", files)
data_files
```

## str_detect()
* `str_detect()` is a function in the tidyverse package that can tell whether a string contains a certain word or phrase. The function returns:
    + `TRUE` if the word you're looking for is in the string
    + `FALSE` if the word you're looking for is *not* in the string
    
\footnotesize
```{r, results='hide'}
str_detect("property_redfin_01.csv", "property") # this returns TRUE
str_detect("location_redfin_01.csv", "property") # this returns FALSE

str_detect(c("property_redfin_01.csv", 
             "location_redfin_01.csv"),
           "location") # c(FALSE, TRUE)
```

## Using str_detect
* Using what we've learned about `str_detect`, let's create a vector that only includes location files.

```{r}
files[str_detect(files, "location")]
```


## Reading in the data - In-Class Exercise
* Now we have all the tools we need to build our dataset!
* You will need to use for-loops and if-else statements to create two datasets, one dataset that includes all the 'location' data and one dataset that includes all the 'property' data. 
    + Initialize two data frames to store all your new data: one data frame for the 'property data' and one for the 'location data'
    + Create a for-loop that cycles through each csv in the folder and checks if it is a 'location' or 'property file'.
    + If it's a property file, add (bind) the csv's data onto your property data frame
    + If it's a location file, bind it into your location_data data frame. 
**Hint(check out what the `bind_rows` function does.)**

\tiny
```{r, eval=FALSE}
property_data <- data_frame() # initialize empty data frame for the property data
 <- data_frame() # initialize empty data frame for the location data

for(file in        ){
    # You only want to have property data in the first dataset
    if(str_detect(     ,"property")){ 
         <- read_csv(   )
        property_data <- bind_rows(property_data, )
    } 
  
  # Now let's combine location data into a different dataset
  else if(str_detect(      , "location")){ 

    
    }
}
```

## Reading in the data - In-Class Exercise (Answer)
\scriptsize
```{r echo=FALSE, message=FALSE}
property_data <- data_frame() # initialize empty data frame
location_data <- data_frame() # initialize empty data frame

for(file in data_files){
# this is a property file, add to property_data
    if(str_detect(file, "property")){ 
        data <- read_csv(file)
        property_data <- bind_rows(property_data, data)
    } 
  # location file
  else if(str_detect(file, "location")){ 
        data <- read_csv(file)
        location_data <- bind_rows(location_data, data)
    }
}
```


## Overview of the Data
* Let's take a look at the variables in our two dataframes - property data and location data
* *What does the property data do? What does the location data do?*
* The good news is that we have much of the data we need, the bad news is that the data is split between location information and price information
* *What piece of information is available in both datasets?*
* We want to combine these two datasets around that common piece of information

## Comparing our Datasets

* Harnessing the power of the `%in%` function

\scriptsize
```{r}
vec1 <- c("green", "yellow", "dog", "teal", "violet", "beige", "cat", "red", "horse", "orange")
vec2 <- c("yellow", "red", "blue", "purple", "green", "orange")
# What does the %in% function show us?
vec1 %in% vec2

# Where are the words that overlap across vectors?
# Where are the words in vector 1 that appear in vector 2?
which(vec1 %in% vec2) 

# How do we return the names of the elements that appear in both vectors?
vec1[vec1 %in% vec2] 
```

## Where to Combine the Data

* Using our %in% function, let's find what variable(s) occur(s) in *both* our property and location datasets.
* We'll save the result as 'overlap', which is a variable we can use for our next step.

```{r, echo = TRUE}
property_names <- names(property_data)
location_names <- names(location_data)

# What variable exists in both datasets?
property_names %in% location_names

overlap <- property_names[property_names %in% location_names]
overlap


```

## "Joining" the Data

* Our URL column exists in both of the dataframes!
* We only want houses that exist in both datasets
* Combining data frames is called "joining." There are multiple types of joins, as you can see in the dplyr cheat sheet:
* Help -> cheat sheets -> Data transformation with dplyr.

\centerline{\includegraphics[height=2.7in]{./Reference/joins.png}}

* Which type of joining would we need to combine our dataframes so that we only have properties with both location and property information?


## "Joining" the Data

\footnotesize
```{r}
joined_data <- inner_join(property_data,
                         location_data,
                         by = overlap)

names(joined_data)
```

## "Joining" the Data

* Success! Now we have one dataset with all of our housing data
* Note - It is also possible to perform a `join` where the columns we are matching in each table do not have the same name using the 'by' argument: by = c("name1" = "name2").
* Similarly, it is also possible to `join` based on multiple columns: by = c(column1, column2, etc...)

## Examining the Data

* Now we can examine the data to check for any problems.
* What are some problems you see in terms of usability for this data?
```{r include = F, eval = F}
View(joined_data)
```

## Examining the Data

### Lots of Issues Here

* price column - USD in front of the numbers
* propertytype column - has "propertytype" in front of every observation
* city_state column - state is inconsistent (va, VA, Virginia) and unwanted states
* lat_lon column - & in the middle of the string
* zip_code column - extra zeros
* sold.dates, list.date columns - dates are not consistent

## Cleaning the Data
* If we try to make a plot without cleaning up our data...it doesn't work too well

```{r echo= FALSE}
joined_data %>% 
    ggplot(aes(x = SQUARE.FEET, y = PRICE)) +
    geom_point()
```

## Cleaning the Data: Price Column
* Let's start by looking at our price data more closely
\footnotesize
```{r}
head(joined_data$PRICE)
```

* There's a USD in front of each number of the column
* What class of variable is the price column?

## Cleaning the Data: Price Column

* How can we fix this problem for a character variable? What type of method could we use?

## Cleaning the Data - In-Class Exercise

* One way we could eliminate the USD is by replacing the USD with nothing, "" (this would be the same as removing it).
* Try using the str_replace function to modify the string "USD40000" to be "40000"

```{r eval = T, echo = T}
str_replace("USD40000", "USD", "")
```

## Cleaning the Data

* Let's mutate the price column to remove "USD" from the character string

\scriptsize
```{r eval = T, echo = T, results = F}
joined_data <- joined_data %>% 
    mutate(PRICE = str_replace(PRICE, "USD", ""))
head(joined_data$PRICE)
```

## Cleaning the Data

* Let's redo the plot!
```{r, echo= FALSE, fig.width=4.5, fig.height=2.5}
joined_data %>% 
    ggplot(aes(x = SQUARE.FEET, y = PRICE)) +
    geom_point() +
    labs(title = "Relationship between price and home size",
         y = "USD",
         x = "Sq ft",
         caption = "Data from Redfin")
```

## Cleaning the Data - In-Class Exercise
* We still have a problem, we forgot to convert our PRICE column to be class numeric!
* We also currently have a warning about missing values -- what should we do about those?
* You should: 
    + Convert the price column to be numeric
    + Remove any missing values from the price and square feet columns
    + Then replot and see how it looks
  
## Cleaning the Data - In-Class Exercise
```{r, eval= FALSE, fig.width=4.5, fig.height=2.5}
joined_data <- joined_data %>% 
    filter(!is.na(      ), !is.na(         )) %>%
    mutate(PRICE = as.numeric(     ))

# Now make the plot
joined_data %>% 
    ggplot(aes(x =      , y =      )) +
    geom_point() +
    scale_y_continuous("USD", labels = dollar) + 
    labs(title = "Relationship between price and home size",
         y = ,
         x = ,
         caption = "Data from Redfin")
```
    
## Cleaning the Data - In-Class Exercise (Answer)
```{r, eval= FALSE, fig.width=4.5, fig.height=2.5}
joined_data <- joined_data %>% 
    filter(!is.na(PRICE), !is.na(SQUARE.FEET)) %>%
    mutate(PRICE = as.numeric(PRICE))

# Now make the plot
joined_data %>% 
    ggplot(aes(x = SQUARE.FEET, y = PRICE)) +
    geom_point() +
    scale_y_continuous("USD", labels = dollar) + 
    labs(title = "Relationship between price and home size",
         y = "USD",
         x = "Sq ft",
         caption = "Data from Redfin")
```

## Cleaning the Data
```{r, echo= FALSE, fig.width=4.5, fig.height=2.5}
joined_data <- joined_data %>% 
    filter(!is.na(PRICE), !is.na(SQUARE.FEET)) %>%
    mutate(PRICE = as.numeric(PRICE))

# Now make the plot
joined_data %>% 
    ggplot(aes(x = SQUARE.FEET, y = PRICE)) +
    geom_point() +
    scale_y_continuous("USD", labels = dollar) + 
    labs(title = "Relationship between price and home size",
         y = "USD",
         x = "Sq ft",
         caption = "Data from Redfin")
```
* What is the relationship between Sq Ft and home price?


## Understanding the Data

* What about the relationship between property type (apt, single family house, condo, etc.) and price?
* Let's produce a table and a bar plot of the average price per square foot by property type.


## Cleaning the Data: Property Type - In-Class Exercise

* Let's clean the property type column
* Check out the PROPERTY.TYPE column and use `str_replace` to fix it

```{r, eval = F}
joined_data <-         %>% 
    mutate(PROPERTY.TYPE = str_replace(          ,
    "propertytype:",   ))
```

## Cleaning the Data: Property Type - In-Class Exercise (Answer)


```{r "in-class exercise answer", echo = F}
joined_data <- joined_data %>% 
    mutate(PROPERTY.TYPE = str_replace(PROPERTY.TYPE,
    "propertytype:", ""))
```


## Understanding the Data
* Let's create a table of average price per square foot by property type
```{r}
type_price <- joined_data %>% 
    group_by(PROPERTY.TYPE) %>% 
    summarise(price = mean(PRICE/SQUARE.FEET, na.rm = T))

type_price
```

## Understanding the Data
Now let's check out a graph. This is obviously not a production-quality graph, just something for our reference.
* How are the colors defined in the graph?

```{r, echo = FALSE}
type_price %>% 
            ggplot(aes(x = PROPERTY.TYPE, 
                       y = price, 
                       fill = PROPERTY.TYPE)) +
            geom_bar(stat = "identity") +
            scale_y_continuous("USD", labels = dollar) +
            scale_fill_discrete(guide = F)  + # get rid of the fill legend
            theme(axis.text.x=element_text(angle=270,hjust=1,vjust=0.5))
```

## Understanding the Data

* City/state is another variable that might impact home sale price... Why?

```{r}
unique(joined_data$CITY_STATE)
```
* What issues do we have with the city_state variable?

## Cleaning the Data

* We can use the `str_split_fixed` function to alter our string (or vector of strings) using a 'pattern'
* The function returns the peices of the string after splitting based on your 'pattern', which could be a word, symbol, or number of characters.

```{r}
# We have three arguments, what do they do?
strings <- str_split_fixed("United States of America", "of", n = 2)

strings

class(strings)
```

## Cleaning the Data

* We told R to split into two pieces (n = 2), so we got two columns in return.
* If we want to get back to individual elements, we can subset the matrix using brackets.

```{r}
strings[1, 1] # first row, first column
strings[1, 2] # first row, second column
```

## Cleaning the Data

* Let's test it out with something closer to our actual use case.
* Note that we're splitting along the comma and a space, which will remove both!

```{r}
str_split_fixed("Arlington, Virginia", ", ", n = 2)
```

## Cleaning the Data - In-Class Exercise

* Use what we just learned to mutate the city_state column and create separate city and state columns.

```{r, eval = FALSE}
joined_data <-       %>% 
    mutate(CITY = str_split_fixed(CITY_STATE,            )[1, 1],
           STATE =                (                       )[1, 2])

joined_data %>% select(CITY_STATE, CITY, STATE) %>% head
```

## Cleaning the Data- In-Class Exercise Answer
```{r, echo = FALSE}
joined_data <- joined_data %>% 
    mutate(CITY = str_split_fixed(CITY_STATE, ", ", n = 2)[,1],
           STATE = str_split_fixed(CITY_STATE, ", ", n = 2)[, 2])

joined_data %>% select(CITY_STATE, CITY, STATE) %>% head
```

## Cleaning the Data: State Data

* We're not done yet. Let's take another look at the state column and see what types of responses exist

\scriptsize
```{r}
unique(joined_data$STATE)
```

* What do you think we need to do to further clean this column?

## Cleaning the Data: State Data

* The function `str_to_upper` will convert all lower case letter to upper case ones
```{r}
str_to_upper("va")
```

* Now use this function to mutate the state column to upper case

## Cleaning the Data: State Data
\scriptsize
```{r, eval = FALSE}
joined_data <-            %>% 
    mutate(STATE =           (    ))

unique(joined_data$STATE)
```

## Cleaning the Data
\scriptsize
```{r}
joined_data <- joined_data %>% 
    mutate(STATE = str_to_upper(STATE))

unique(joined_data$STATE)
```

## Cleaning the Data - In-Class Exercise
* Now we need to change the instances of "VIRGINIA" to "VA"
* We've already used this function! 
* Go back and find the function you need to make that change

```{r "VIRGINIA ANSWER", echo = F, include = F}
joined_data <- joined_data %>% 
    mutate(STATE = str_replace(STATE, "VIRGINIA", "VA"))
```

## Cleaning the Data - In-Class Exercise

* The last thing we need to do is toss out the rows with MICHIGAN and COLORADO since we only want data along the DC Metro
* Use the `%in%` function to filter only rows with VA, DC, or MD in them.

## Cleaning the Data

```{r}
joined_data <- joined_data %>% 
    filter(STATE %in% c("VA", "DC", "MD"))

unique(joined_data$STATE)
```

## Plotting the Data
* Let's find out the average prices by state for our dataset

```{r, eval = FALSE}
state_average <- joined_data %>% group_by(STATE) %>% 
    summarise(price = mean(PRICE/SQUARE.FEET, na.rm = T))

state_average_plot <- state_average %>% 
    ggplot(aes(x = STATE, y = price, fill = STATE)) +
    geom_bar(stat = "identity", width = 0.5) + 
    labs(x = NULL, # What does this do?
         y = "USD/Sq ft",
         title = "Average price for DC Home sales",
         caption = "Data from Redfin") 

state_average_plot
```

## Plotting the Data
* I want to have a larger plot title and have it centered
* How can we do that? Would we use a data-related dimension or an aesthetic option?

```{r, echo = FALSE}
state_average <- joined_data %>% group_by(STATE) %>% 
    summarise(price = mean(PRICE/SQUARE.FEET, na.rm = T))

state_average_plot <- state_average %>% 
    ggplot(aes(x = STATE, y = price, fill = STATE)) +
    geom_bar(stat = "identity", width = 0.5) + 
    labs(x = NULL, # What does this do?
         y = "USD/Sq ft",
         title = "Average price for DC Home sales",
         caption = "Data from Redfin") 

state_average_plot
```

## Plotting the Data
* We can use the `theme()` function to alter elements of the graph.
* What other textual elements can we alter in this graph?

\scriptsize
```{r, fig.width=4, fig.height=2}
state_average_plot +
    theme(plot.title = 
            element_text(size = 14, hjust = 0.5, color = "Red"))
```

## Plotting the Data

* Since our title, and the other labels, is a character string, or text, we use element_text to control it. element_text allows us to set:

    + font family (Times New Roman, Arial, etc...) 
    + font size (10, 12, 14, etc...)  
    + font face (bold, italic) 
    + color 
    + hjust (horizontal adjustment) 
    + vertical adjustment
    + angle  
    + other aspects that impact the display of text.

## Plotting the Data

\scriptsize
```{r, , eval = FALSE, fig.width=5, fig.height=3}
state_proptype <- joined_data %>% 
    group_by(PROPERTY.TYPE, STATE) %>% 
    summarise(price = mean(PRICE/SQUARE.FEET, na.rm = T)) # empty lots

plot <- state_proptype %>% 
    ggplot(aes(x = STATE, y = price, fill = STATE)) +
    geom_bar(stat = "identity") +
    facet_wrap("PROPERTY.TYPE") +
    labs(title = "Price by property type",
         x = NULL,
         y = "USD/Sq ft.",
         caption = "Data from Redfin")

plot + theme(plot.title = element_text(face = "italic", size = 14, color = "Blue"),
          axis.text.x = element_text(face = "bold", color = "green4"),
          plot.background = element_rect(fill = "beige", color = "red"),
          panel.background = element_rect(fill = "cadetblue1", 
                                          color = "black", linetype = 3))
              
              
```

## Plotting the Data
\normalsize
```{r, echo = FALSE, fig.width=4.5, fig.height=3}
state_proptype <- joined_data %>% 
    group_by(PROPERTY.TYPE, STATE) %>% 
    summarise(price = mean(PRICE/SQUARE.FEET, na.rm = T)) # empty lots
state_proptype %>% 
    ggplot(aes(x = STATE, y = price, fill = STATE)) +
    geom_bar(stat = "identity") +
    facet_wrap("PROPERTY.TYPE") +
    labs(title = "Price by property type",
         x = NULL,
         y = "USD/Sq ft.",
         caption = "Data from Redfin") +
    theme(plot.title = element_text(face = "italic", size = 14, color = "Blue"),
          axis.text.x = element_text(face = "bold", color = "green4"),
          plot.background = element_rect(fill = "beige", color = "red"),
          panel.background = element_rect(fill = "cadetblue1", 
                                          color = "black", linetype = 3))
              
              
```

## Understanding the Data

* We used `element_rect()` to adjust rectangular elements of the plot.
* The main plot has a beige background with a red border.
* We used the panel.background option to adjust elements of the area where the data is
* What did the linetype argument do in our panel.background option?
* What other rectangles could we adjust?

```{r, echo= FALSE, eval = FALSE, fig.width=5, fig.height=3}
state_proptype <- joined_data %>% 
    group_by(PROPERTY.TYPE, STATE) %>% 
    summarise(price = mean(PRICE/SQUARE.FEET, na.rm = T)) # empty lots

state_proptype %>% 
    ggplot(aes(x = STATE, y = price, fill = STATE)) +
    geom_bar(stat = "identity") +
    facet_wrap("PROPERTY.TYPE") +
    labs(title = "Price by property type",
         x = NULL,
         y = "USD/Sq ft.",
         caption = "Data from Redfin") +
    theme(plot.title = element_text(face = "italic", size = 14, color = "Blue"),
          axis.text.x = element_text(face = "bold", color = "green4"),
          plot.background = element_rect(fill = "beige", color = "red"),
          panel.background = element_rect(fill = "cadetblue1", 
                                          color = "black", linetype = 3))
              
              
```

## Understanding the Data
* Perhaps using the price per square foot is not the best method of analysis.
* When a property is sold the surrounding land around the house will likely affect the selling price.
* Let's work on making a plot of price/lot size

## Understanding the Data
* We'll start by looking at the head of the data
```{r, echo = FALSE}
joined_data %>% select(PRICE, SQUARE.FEET, LOT.SIZE) %>% head()
```
* We have a lot of NA values. Why is that?

## Understanding the Data - In-Class Exercise

* Let's mutate the lot.size column so that if there is an `NA` then we replace it with the value of the square.feet column and otherwise keep lot.size the same

## Understanding the Data

\scriptsize
```{r, eval = FALSE}
joined_data2 <- joined_data %>% 
    mutate(LOT.SIZE = ifelse(is.na(LOT.SIZE), SQUARE.FEET, LOT.SIZE))

# Now we can find our average
lot_state_data <- joined_data %>% 
    group_by(PROPERTY.TYPE, STATE) %>% 
    summarise(price = mean(PRICE/LOT.SIZE, na.rm = T))

head(lot_state_data)
```

## Understanding the Data
```{r, echo = FALSE}
joined_data <- joined_data %>% 
    mutate(LOT.SIZE = ifelse(is.na(LOT.SIZE), SQUARE.FEET, LOT.SIZE))

# Now we can find our average
lot_state_data <- joined_data %>% 
    group_by(PROPERTY.TYPE, STATE) %>% 
    summarise(price = mean(PRICE/LOT.SIZE, na.rm = T))

head(lot_state_data)
```

## Understanding the Data
* The culminating exercise of the day!
* Make a plot like the one we just did but using our new price/lot.size variable with some modifications -
    + **Make your title centered and Green**
    + **Make your panel background White with a red border**

```{r "Lot size plot answer", echo = F, include = F}
lot_state_data %>% 
    ggplot(aes(x = STATE, y = price, fill = STATE)) +
    facet_wrap("PROPERTY.TYPE") +
    geom_bar(stat = "identity", width = 0.5) +
    labs(x = NULL,
         title = "Price per square foot of total property size",
         caption = "Data from Redfin",
         y = "USD/Sq ft.") +
    theme(plot.title = element_text(hjust = 0.5, color = "green"),
          panel.background = element_rect(fill = "white", color = "red"))
```

## Understanding the Data

```{r, echo = F, fig.width=4.5, fig.height=3}
lot_state_data %>% 
    ggplot(aes(x = STATE, y = price, fill = STATE)) +
    facet_wrap("PROPERTY.TYPE") +
    geom_bar(stat = "identity", width = 0.5) +
    labs(x = NULL,
         title = "Price per square foot of total property size",
         caption = "Data from Redfin",
         y = "USD/Sq ft.") +
    theme(plot.title = element_text(hjust = 0.5, color = "green"),
          panel.background = element_rect(fill = "white", color = "red"))
```