---
title: "Module 1 Day 3"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

```{r}
# install.packages("plyr")
library(plyr)
library(dplyr)
library(ggplot2)
# install.packages("stargazer")
library(stargazer)
# install.packages("Hmisc")
library(Hmisc)
# install.packages("lubridate")
library(lubridate)
# install.packages("ggthemes")
library(ggthemes)
```

## Recap last week

Set your working directory

```{r}
setwd()
```

* Read in the data file "fred_median_income.csv"
* Convert the data column from a character to a Date
* Filter to data post 2005
* Make a scatter plot  with the date on the x axis and median_hh_income on the y axis
* Be sure to label your chart appropriately
* Use the scale_x_dates() function to label every 2 years

```{r}
fred_median_income <- 
  read_csv()

fred_median_income %>%
  mutate(date = ) %>%
  filter() %>%
  ggplot() +
  geom_point() +
  labs(x = "Date", 
       y = "Median Household Income",
       title = "Real Median Household Income over Time",
       caption = "Source: FRED") +
  scale_x_date() 

```

What do we mean when we say "real" income?

## Shrinking Real Incomes

* Use the lag() function to calculate the percent change in household income
* Add a new column called shrunk that "Yes" if the percent change is less than zero and "No" otherwise
* Make a bar chart of the percent change in median household income since 2000
* Use the variable shrunk as the fill axis
* Use geom_hline() to add a horizontal line at y = 0 to highlight years where real income shrunk
* Turn off the fill axis by using guides(fill = FALSE)
* Use scale fill manual so that when income shrinks the point is red


```{r, fig.width=6, fig.height=3.5, echo=FALSE}
fred_median_income %>%
  mutate(date = ),
         median_hh_income_chng = ,
         shrunk = ifelse( , , )) %>%
  filter(date >= ) %>%
  ggplot() +
  geom_col() +
  labs(x = "Date", 
       y = "Percent Change",
       title = "Changes Real Median Household Income over Time",
       caption = "Source: FRED") +
  scale_x_date() +
  geom_hline(yintercept = ) +
  guides(fill = FALSE) +
  scale_fill_manual(values = c())

```

## Recap last week

* Read in the data file "acs_2016_sample.csv"
* Apply our standard transformations
* Create a new data frame called acs_mean_median that is the weighted mean and median income by race and sex
* Add a column that is the difference between the mean and median income 
* Which group has the biggest difference? Which the smallest?

```{r}
acs_2016_transformed <- 
  read.csv("acs_2016_sample.csv") %>%
  rename(state_code = STATEFIP,
         sex = SEX,
         age = AGE,
         race = RACE,
         hispanic = HISPAN,
         education = EDUC,
         employment_status = EMPSTAT,
         total_income = INCTOT,
         wage_income = INCWAGE,
         hrs_worked = UHRSWORK,
         weeks_worked = WKSWORK2,
         weight = PERWT,
         metro = METRO) %>%
  select(state_code, sex, age, race, hispanic, education, 
         employment_status, total_income, wage_income, 
         hrs_worked, weeks_worked, weight, metro) %>%
  mutate(sex = ifelse(sex == 1, "Male", "Female"),
         education = case_when(education <= 5 ~ "No HS Diploma",
                               education == 6 ~ "HS Diploma",
                               education %in% c(7,8,9) ~ "Some College",
                               education == 10 ~ "College Degree",
                               education == 11 ~ "Graduate Degree"),
         race = case_when(race == 1 & hispanic == 0 ~ "Non-hispanic White",
                          race == 1 & (hispanic %in% c(1,2,3,4)) ~ "Hispanic White",
                          race == 2 ~ "Black",
                          race == 3 ~ "Native American",
                          race %in% c(4,5,6) ~ "Asian",
                          race %in% c(7,8,9) ~ "Other/Multiracial"),
        weeks_worked = case_when(weeks_worked == 1 ~ 13,
                                 weeks_worked == 2 ~ 26,
                                 weeks_worked == 3 ~ 39,
                                 weeks_worked == 4 ~ 47,
                                 weeks_worked == 5 ~ 49,
                                 weeks_worked == 6 ~ 52)) %>%
  filter(total_income < 9999999)
```


```{r}
acs_mean_median <-
  
```

* Going back to the whole sample filter to only those groups
* Use the paste function to create a new variable called group that is paste(sex, race)
* Make a density plot of the income of the two groups
* Set the alpha parameter to 0.5

## Recap last week

```{r}
acs_2016_transformed %>%
  filter(() | ()) %>%
  mutate(group = paste()) %>%
  ggplot() +
  geom_density() +
  labs(x = "Total Income",
       y = "Density",
       color = "",
       fill  = "",
       caption = "Source: 2016 ACS")
```
What does it mean for the distribution if there is a big difference between the mean and median?

## Prepping the data

Create a new data frame acs_2016_cleaned which is acs_2016_transformed filtered to people:

* Between 18 and 65
* In the workforce
* With a total wage <= 1,000,000
* Worked more than 0 hours a week
* Worked more than 0 weeks

* Add a column for hourly wage

What fraction of the original observations do we have?

```{r}
acs_2016_cleaned <-

```

## Preping the data

Select the columns 
 
 * wage_income, age, hrs_worked , weeks_worked, and hourly_wage
 
 Make a stargazer summary table

```{r, results='asis'}
stargazer(acs_2016_cleaned %>%
          select(),
          header = F, rownames = F, title = "Summary Statistics", omit.summary.stat = "n")
```


## OLS Regression 

* Try it out! Run a simple regression of salary on ages 

```{r}
baseline_model <- lm()
```


How do we add weights?

```{r}
baseline_model <- lm(, weights = weight,
                     acs_2016_cleaned)
```

## stargazer()

* Once again, we use stargazer to look at the results

```{r}

```


## The Broom Package

* Let's try it out!
* tidy, augment, and glance at the results of the baseline model
* How can we use the augment function to keep all of our original columns?

```{r}
tidy()

glance()


baseline_model_augmented <- augment(baseline_model, acs_2016_cleaned)
```

## Improving our model

* Let's make a plot of the distribution of residuals by gender.
* What do we learn from this chart?

```{r}
ggplot() +
geom_boxplot(aes(x = , y = .resid)) +
ylim(c(-10,10)) +
labs(x = "Residual",
     y = "Density",
     title = "Distribution Residuals by Gender",
     color = "") 
```

## Improving our model
* Let's update the code for our model:

```{r}
improved_model <- lm(hourly_wage ~ age + sex, 
                     weights = weight,
                     acs_2016_cleaned)
```

* It's real easy to compare multiple models with stargazer()
```{r, results='asis'}
stargazer(baseline_model, improved_model)
```

## Put my models to shame

* Pair up!
* Take 15 - 20 mins to improve on the models we have done so far.
* I want to see plots that explain why you are adding in variables
* I want to see beautiful regression output tables
* I want you to spend 5 minutes writing up a post on piazza that includes a graph, a table, and a brief explanation of your model
  