---
title: "Module 2"
output:
  beamer_presentation: default
  ioslides_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, size = "tiny", width = 50, fig.height = 3.5,
                      fig.width = 5)
library(tidyverse)
library(stargazer)
```

# Day 3

## Recap last week
* "Joined" location and property data using the url column
* Introduce `str_replace`, `str_split_fixed`, `str_trim`, and `str_to_upper` to clean variables in our data


## Goals for Today

R:

* Finish cleaning current data
* Motivate the use of user-created functions and "function safety"
* Learn how to create functions
* Create our own function to find the distance from each house to the nearest house

Economics:

* Take 
* Begin exploring the relationshiop between distance from a metro stop and home prices

## Cleaning zip-codes

* Let's investigate our zip-code variable

\scriptsize
```{r, message=FALSE,eval=FALSE}
dat_path <- ## Your file path here
joined_data <- read_csv(paste0(dat_path,
                               "joined_data.csv"))
head(joined_data$ZIP_CODE)
```

```{r, message=FALSE,echo=FALSE}
dat_path <- "~/Class_Materials/Spring_2018/module_2/day_3/"
joined_data <- read_csv(paste0(dat_path,
                               "joined_data.csv"))
head(joined_data)
```

* We can see two big problems with our current data:
    + Each zip-code has 4 extra zeros
    + There are quotation marks around each observation

## str_sub and str_length

* `str_length()` tells us the number of characters in a string
* `str_sub()` returns part of a text string between the start and end position provided
```{r}
str_length("Hello")
str_sub("Hello", start = -1, end = 3)
```

## str_sub and str_length

* `str_length()` is primarly used with other functions
* We can combine both functions in order to extract the actual zip-code from our variable

```{r}
str_sub("'222040000", 2, 
        str_length("'222040000") - 4)
```

* **Use dplyr and these new stringr functions to fix the rest of our zip-code data**

```{r, echo = FALSE}
joined_data <- joined_data %>% 
    mutate(ZIP_CODE = str_sub(ZIP_CODE, 2,
                      str_length(ZIP_CODE) - 4))
```

## Reviewing the Data

* We'll start by looking at the head of the data
```{r, echo = FALSE}
joined_data %>% select(PRICE, SQUARE.FEET, LOT.SIZE) %>% head()
```
* We have a lot of NA values. Why is that?

## Cleaning The Data
* Let's get rid of the `NA` values by replacing them with square footage
* We'll create a new column called Property Size that equals either lot size or square feet
      + Use an `ifelse()` statement in our mutate, where we test whether lot size is `NA`
      + If lot size is `NA`, use square feet. Otherwise, use lot size
* Next we will summarize our data by price per foot of property size. 
      + We want to group by property type and state!

## Cleaning Code
\scriptsize
```{r}
joined_data <- joined_data %>% 
    mutate(PROPERTY.SIZE = ifelse(is.na(LOT.SIZE), SQUARE.FEET, LOT.SIZE))

# Now we can find our average
lot_state_data <- joined_data %>% 
    group_by(PROPERTY.TYPE, STATE) %>% 
    summarise(price = mean(PRICE/PROPERTY.SIZE, na.rm = TRUE))

head(lot_state_data)
```


## Visualizing The Relationship
* **Make a plot showing the average price per lot size for each property type in each state**

```{r "Lot size plot answer", eval = FALSE, include = FALSE}
lot_state_data %>% 
    ggplot(aes(x =      , y =     , fill =     )) +
    facet_wrap("PROPERTY.TYPE") +
    geom_bar(stat = "identity", width = 0.5) +
    labs(x = NULL,
         title = "      ",
         caption = "           ",
         y = "         ")
```

## Finished Plot
```{r, echo = FALSE, fig.width=4.5, fig.height=3}
lot_state_data %>% 
    ggplot(aes(x = STATE, y = price, fill = STATE)) +
    facet_wrap("PROPERTY.TYPE") +
    geom_bar(stat = "identity", width = 0.5) +
    labs(x = NULL,
         title = "Price per square foot of total property size",
         caption = "Data from Redfin",
         y = "USD/Sq ft.")
```

## Different Forms of Data
* There are two types of data that we work with:
    1. Long
    2. Wide
* Up until now we have been looking at only one format of data --- long type
    + Much easier to manipulate and graph
* Wide-form data is easier to see visually
    + Generally used to display tables
    
## A Real Estate Example
* Let's look at a toy example with data from Redfin
```{r, warning=FALSE, message=FALSE}
wide_data <- read_csv(paste0(dat_path,
                               "wide_sample.csv"))
```

* Each observation is a property at every point in time for our data
    + How is this different from an observation in our original data?
    + If this data were in long form, how many columns would there be?

## Converting Our Data
\centerline{\includegraphics[height=3in]{spread_gather.png}}

## gather()
* `gather(data, key, value, ...)`
    + Converts multiple columns from wide-format to long-format
    + `key` - name of the variable you are trying to create
    + `value` - variable that holds the values in the key variables
    + Must specify which columns to use for the key
```{r}
long_data <- wide_data %>% 
  gather(key = "year", value = "price", 5:11)
```

## spread()
* `spread(data, key, value, ...)`
    + Converts two columns from long-format to wide-format
    + `key` - variable you wish to spread into multiple columns
    + `value` - variable that holds the values that will be used by your new variables
```{r}
new_wide <- long_data %>%
  spread(key = "year", value = "price")
```

## Other Uses
* We can use our reshaping functions in conjunction with stargazer to produce good looking tables

\scriptsize
```{r, results='asis'}
lot_state_data %>% 
  spread(key = STATE, value = price) %>%
  stargazer(summary = FALSE, 
            header = FALSE,
            type = "text")
```
## Reading In The Metro Station Data
* Use this data in conjunction with the housing data to find the distance from a metro
    + Includes latitude and longitude of each metro station
* Unfortunately it is in an xlsx file so we can't use read_csv
```{r, message=FALSE}
metro_data <- read_xlsx(paste0(dat_path,
                        "Metro_lat_lon.xlsx"), 
                        sheet = 1)
```
* There is one big problem with finding the distance --- the calculation is fairly difficult
    + We need to create a function!

## What Is A Function
* Piece of code that takes one or more inputs and returns one or more outputs
* For example:
    +  `min()` takes a vector of numbers and returns the number with the lowest value
    +  `read_csv()` takes a string that contains a file path and returns a data.frame object
* Some of inputs may have a default set value for certain arguments
    + `min(..., na.rm = FALSE)`
    
## Writing A Function
* Functions are extremely useful when you have to repeat a complex chunk of code
    + Can be used in loops, if else statements, and even other functions
* A user-defined function takes the following form:
```{r, eval = FALSE}
function_name <- function(arguments){
  ## Code that does your task
}
```


## Simple Examples
* Here is a funciton for adding two numbers together:
```{r}
add <- function(num1, num2){
    num1 + num2
}
add(num1 = 2,num2 = 5)
```

* **Create your own function that raises a number provided by the user to a power provided by the user. **

```{r "In-class 2 Answer", echo=FALSE}
exponent <- function(base, power){
    base**power
}
```

## Defaults
* Use these when one option is used much more than any of the other options
* For example `read_csv(..., col_names = TRUE)`
    + read in the first row of the data as column names
    + may also use `FALSE` or a vector of column names we would like to assign
* Merely set one of our arguments equal to the default value
```{r, eval = FALSE}
root <- function(base, denom = 2){
  base**(1/denom)
}
```

## return()
* Let's take a look the functions we created so far
    + What do we input and what do we output?
    + Is this completely clear from our code?
* By default, a function returns the last object that was created
    + Can be confusing when reading someone elses code
* `return()` says I want the output of my function to be this value
    + Only useable inside of a user-generated function
    
## A More Complex Example
* We can use any of the coding structures we've already learned inside of our functions
* Let's create a function that sums together the numbers in a vector
```{r}
vector_sum <- function(numeric_vector){
    sum <- 0
    for(number in numeric_vector){
        sum <- sum + number
    }
    return(sum)
}
vector_sum(c(1,1,5,12))
```

## Another Complex Example
* Let's find the minimum number in a vector
```{r}
vector_min <- function(numeric_vector){
    m <- numeric_vector[1]
    for(number in 2:length(numeric_vector)){ 
        if(numeric_vector[number] < m){
            m <- numeric_vector[number]
        }
    }
    return(m)
}

test_vector <- c(1, 4, 6, -8, 0, 11)
vector_min(test_vector)
```

## Another Complex Example

**Now it's your turn. Write a function called vector_max which takes a numeric vector and returns the maximum element in it. Make sure to test out your function. **
```{r "vector_max answer", echo = FALSE, include = FALSE}
vector_max <- function(numeric_vector){
    
    m <- numeric_vector[1] # initialize our minimum to the first element
    
    for(i in 2:length(numeric_vector)){ 
        # Now go through the rest of the vector
        if(numeric_vector[i] > m){
            m <- numeric_vector[i]
        }
    }
    return(m)
}
test_vector <- c(1, 4, 6, -8, 0, 11)
vector_max(test_vector)
```

## Rock, Paper, Scissors
* Let's try using a more fun example now with a simple logic game
```{r, eval = FALSE}

RPS <- function(player1, player2){
if(( player1 == "Rock" & player2 == "Scissors") | ( player1 == "Paper" & player2 == "Rock") | ( player1 == "Scissors" & player2 == "Paper")){
        return("Player 1 Wins!")
} else if(player1 == player2){
    return("Tie game!")
} else {
    return("Player 2 Wins!")
}
}

test <- RPS("Scissors", "Rock")

test
```

**Fill in the logical arguments above. Create a function called RPS that uses the code as the body of the function and returns the results of the match.**

## Returning to Metro Data
* Remember, our goal is to calculate the nearest metro station
    + Use our longitude and latitude data combined with a little trigonometry to find the distance
    + Remember: $a^2 + b^2 = c^2$ or $\Delta longitude^2 + \Delta latitude^2 = distance^2$

## Calculating Distance

* Now that our data is numeric we can calculate the distance

\scriptsize
```{r}
test_property <- joined_data[1, ]
test_metro <- metro_data[1, ]

delta_x <- test_property$LONGITUDE - test_metro$Longitude
delta_y <- test_property$LATITUDE - test_metro$Latitude

distance <- sqrt(delta_x**2 + delta_y**2)
# One degree is equal to about 69 miles
distance/69
```

## Basic Distance Function

**Write a function that takes 4 arguments: property long and lat values, and metro long and lat values. The function should then calculate the pythagorean distance between the property and the metro in miles.**

```{r, eval=FALSE}
metro_dist <- function( , , , ){
  
  
}
```

```{r, echo = FALSE}
metro_dist <- function(prop_long, prop_lat, metro_long, metro_lat){
  delta_x <-prop_long - metro_long
  delta_y <- prop_lat - metro_lat

  dist <- sqrt(delta_x**2 + delta_y**2)
  miles <- dist/69
  
  return(miles)
}
```

## Double checking the distance
* Our test property is in Arlington and our test metro is in Maryland east of D.C.
    + That's about 10 miles
    + Why might our distance be incorrect?
    
## geosphere
* We can't use the simple Pythagorean formula
* `geosphere` is a package that allows us to easily work with geographic coordinates
* `distHaversine()` calculates the shortest distance between two points
    + Points must contain longitude AND latitude
    + "As the crow flys"
    + Default output is in meters
```{r, message=FALSE}
# install.packages("geosphere")
library(geosphere)
?distHaversine()
```

## Calculating distance
```{r}
prop_long_lat <- c(test_property$LONGITUDE,
                   test_property$LATITUDE)
metro_long_lat <- c(test_metro$Longitude,
                    test_metro$Latitude)

distance <- distHaversine(prop_long_lat, metro_long_lat)
# There are 1609.344 meters in a mile
distance/1609.344
```

* This looks much more realistic!

## Every Metro Stop
\scriptsize
```{r, results='hide'}
output_distances <- c()

for(i in 1:nrow(metro_data)){
    # Pull out our longitude and latitude values for the metro station
    metro_long_lat <- c(metro_data$Longitude[i],
                        metro_data$Latitude[i])
    # Calculate the distance to our property
    distance <- distHaversine(p1 = prop_long_lat, # from above
                              p2 = metro_long_lat)
    # append that distance to our distances vector
    output_distances <- c(output_distances, distance)
}

output_distances
```

## Better Distance Function
**Turn the code from the previous slide into a function called prop_metros_dist that does the following:**

* **Finds the distance from a property to every metro station**
* **Converts the distance from meters to miles**
* **Returns the distance (in miles) of the closest metro station**

```{r "prop_metro_answer", echo = FALSE, results='hide'}
prop_metros_dist <- function(prop_long_lat, metro_data){
    output_distances <- c()

    for(metro in 1:nrow(metro_data)){
        # Pull out our longitude and latitude values for the metro station
        metro_long_lat <- c(metro_data$Longitude[metro],
                            metro_data$Latitude[metro])
        # Calculate the distance to our property
        distance <- distHaversine(p1 = prop_long_lat,
                                  p2 = metro_long_lat)
        # append that distance to our distances vector
        output_distances <- c(output_distances, distance)
    }
    return(min(output_distances)/1609.344)
}

prop_metros_dist(prop_long_lat, metro_data)
```

## Function safety
* Functions help use improve readability and replicability of our code, but are not fool proof
* People often input the wrong arguments or use functions in ways that weren't intended
* Let's look at `prop_metros_dist()`:
    + If a user doesn't recoginze `distHaversine()` is from the geosphere package, they may not install and load it
    + It is our job as the programmer to address these types of concerns
    
## Functions for Function safety
* `require()` --- works similarly to `library()`
    + Designed to be used within other functions
    + Returns `FALSE` if a package fales to load
    + Can use this with if/else statements
* `warning()` --- prints a warning message that we specify but DOES NOT stop the program
* `stop()` --- prints an error message and stops the program

## Function safety in Action
\scriptsize
```{r}
prop_metros_dist <- function(prop_long_lat, metro_data){
  
  if(require(geosphere) != TRUE){ # what does require do?
      warning("geosphere package is not installed")
    } else{  
  output_distances <- c()
  for(i in 1:nrow(metro_data)){
        metro_long_lat <- c(metro_data$Longitude[i],
                            metro_data$Latitude[i])
        distance <- distHaversine(p1 = prop_long_lat,
                                  p2 = metro_long_lat)
        output_distances <- c(output_distances, distance)
    }

    return(min(output_distances)/1609.344)
    }
}

```

## Wrapper Functions
* Need a two-number vector with longitude and latitude but we have two columns in our data
* A function that calls another function
    + Generally pre-process inputs for other functions
    + Improve readibility, ease of use, and our ability to implement changes
    + `prop_metros_dist()` calls `dist_Haversine()`
* Want to write a function that turns `LATITUDE` and `LONGITUDE` into a vector

## Our Wrapper Function
\scriptsize
```{r}
distance_function <- function(LONG, LAT, metro_df){
    if(!require(geosphere)){ # what does require do?
      warning("geosphere package is not installed")
    } else{
      # We need to turn our LONG and LAT columns into a vector
      prop_long_lat <- c(LONG, LAT)
      # Now call our prop_metros_dist function from above
      out <- prop_metros_dist(prop_long_lat,
                            metro_df)
      return(out)
    }
}

distance_function(joined_data$LONGITUDE[3],
                  joined_data$LATITUDE[3],
                  metro_data) 
```

## Finding the Closest Metro Station
* **Combine our function with a loop to find the closest metro station to each property (Hint: Look at one of our previous functions) **
```{r, eval = FALSE}
joined_data$metro_distance <- NA #initialize the column
for(){

  ## Your code here
  
}


```

```{r,echo = FALSE, results = 'hide'}
joined_data$metro_distance <- NA #initialize the column

for(i in 1:nrow(joined_data)){
    joined_data$metro_distance[i] <- 
      distance_function(joined_data$LONGITUDE[i],                                                   joined_data$LATITUDE[i],
                        metro_data)
}
head(joined_data$metro_distance)
```


## Graphing the Relationship

```{r, echo = FALSE, eval = FALSE}
joined_data <- joined_data %>% 
    mutate(distance_tenth = round(metro_distance, 0.1))

joined_data %>% group_by(distance_tenth) %>% 
    summarize(price = mean(PRICE/SQUARE.FEET, na.rm = T)) %>% 
    ggplot(aes(x = distance_tenth, y = price)) +
    geom_line() +
    labs(title = "Price vs distance from metro station",
         y = "USD/Sq ft.",
         x = "Miles",
         caption = "Data from Redfin and WMATA")
```