---
title: "Day 3 Student"
author: "Ian de Medeiros"
date: "March 22, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(stargazer)
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```


## Cleaning the zip codes
```{r}
dat_path <- ## Your file path here
  joined_data <- read_csv(paste0(dat_path,
                                 "joined_data.csv"))
head(joined_data$ZIP_CODE)
```

## str_sub() and str_length()
```{r}
str_length("Hello")
str_sub("Hello", start = 2, end = 4)

str_sub("'222040000", 2, 
        str_length("'222040000") - 4)
```

## In-Class 1:

Use dplyr and these new stringr functions to fix the rest of our zip-code data

```{r}
joined data <- # Your answer here
```

## Lot Size
```{r}
joined_data %>% select(PRICE, SQUARE.FEET, LOT.SIZE) %>% head()

joined_data <- joined_data %>% 
  mutate(PROPERTY.SIZE = ifelse(is.na(LOT.SIZE), SQUARE.FEET, LOT.SIZE))

# Now we can find our average
lot_state_data <- joined_data %>% 
  group_by(PROPERTY.TYPE, STATE) %>% 
  summarise(price = mean(PRICE/PROPERTY.SIZE, na.rm = TRUE))

head(lot_state_data)
```

## In-Class 2:

Make a plot showing the average price per lot size for each property type in each state

```{r}
lot_state_data %>% 
  ggplot(aes(x =      , y =     , fill =     )) +
  facet_wrap("PROPERTY.TYPE") +
  geom_bar(stat = "identity", width = 0.5) +
  labs(x = NULL,
       title = "      ",
       caption = "           ",
       y = "         ")
```

## Wide-format Data
```{r, warning=FALSE, message=FALSE}
wide_data <- read_csv(paste0(dat_path,
                               "wide_sample.csv"))
```

## spread() and gather()
```{r}
long_data <- wide_data %>% 
  gather(key = "year", value = "price", 5:11)

wide_data <- long_data %>%
  spread(key = "year", value = "price")
```

## Other uses
```{r}
spread(key = STATE, value = price) %>%
  stargazer(summary = FALSE, 
            header = FALSE,
            type = "text"")
```

## Metro Data
```{r, message=FALSE}
library(xlsx)
metro_data <- read.xlsx(paste0(dat_path,
                        "Metro_lat_lon.xlsx"), 
                        sheetIndex = 1)
```

## Simple function
```{r}
add <- function(num1, num2){
    num1 + num2
}
add(num1 = 2,num2 = 5)
```

## In-Class 3

Create your own function that raises a number provided by the user to a power provided by the user.

```{r}
exponent <- function( , ){
  # Your answer here
}
```

## More complex examples
```{r, eval = FALSE}
root <- function(base, denom = 2){
  base**(1/denom)
}
```

```{r}
vector_sum <- function(numeric_vector){
    sum <- 0
    for(number in numeric_vector){
        sum <- sum + number
    }
    return(sum)
}
vector_sum(c(1,1,5,12))
```

```{r}
vector_min <- function(numeric_vector){
    m <- numeric_vector[1]
    for(number in 2:length(numeric_vector)){ 
        if(numeric_vector[number] < m){
            m <- numeric_vector[number]
        }
    }
    return(m)
}

test_vector <- c(1, 4, 6, -8, 0, 11)
vector_min(test_vector)
```

## In-Class 4

Now it's your turn. Write a function called vector_max which takes a numeric vector and returns the maximum element in it. Make sure to test out your function.

```{r}
vector_max <- function( ){
  # Your answer here
}
```

## In-Class 5

Fill in the logical arguments above. Create a function called RPS that uses the code as the body of the function and returns the results of the match.

```{r, eval = FALSE}
if(( & ) | ( & ) | ( & )){
        print("Player 1 Wins!")
} else if (){
    print("Tie game!")
} else {
    print("Player 2 Wins!")
}
```

## Calculating Distance
```{r}
delta_x <- test_property$LONGITUDE - test_metro$Longitude
delta_y <- test_property$LATITUDE - test_metro$Latitude

distance <- sqrt(delta_x**2 + delta_y**2)
# One degree is equal to about 69 miles
distance/69
```

## In-Class 6

Write a function that takes 4 arguments: property long and lat values, and metro long and lat values. The function should then calculate the pythagorean distance between the property and the metro in miles.

```{r, eval=FALSE}
metro_dist <- function( , , , ){
  
  
}
```

## Ignoring the Flat Earth Hypothesis
```{r, message=FALSE}
# install.packages("geosphere")
library(geosphere)
?distHaversine()

prop_long_lat <- c(test_property$LONGITUDE,
                   test_property$LATITUDE)
metro_long_lat <- c(test_metro$Longitude,
                    test_metro$Latitude)

distance <- distHaversine(prop_long_lat, metro_long_lat)
# There are 1609.344 meters in a mile
distance/1609.344
```

## Calculating the Distance from Every Metro Stop
```{r}
output_distances <- c()

for(metro in 1:nrow(metro_data)){
    # Pull out our longitude and latitude values for the metro station
    metro_long_lat <- c(metro_data$Longitude[metro],
                        metro_data$Latitude[metro])
    # Calculate the distance to our property
    distance <- distHaversine(p1 = prop_long_lat, # from above
                              p2 = metro_long_lat)
    # append that distance to our distances vector
    output_distances <- c(output_distances, distance)
}

output_distances
```

## In-Class 7

Turn the code from the previous slide into a function called prop_metros_dist that does the following:
* Finds the distance from a property to every metro station
* Converts the distance from meters to miles
* Returns the distance (in miles) of the closest metro station

```{r}
prop_metros_dist <- function( , ){
    output_distances <- c()
    # Your anwser here
}

prop_metros_dist(prop_long_lat, metro_data)
```

## Function Safety

```{r}
prop_metros_dist <- function(prop_long_lat, metro_data){
  
  if(!require(geosphere)){ # what does require do?
      warning("geosphere package is not installed")
    } else{  
  output_distances <- c()
  for(i in 1:nrow(metro_data)){
        metro_long_lat <- c(metro_data$Longitude[i],
                            metro_data$Latitude[i])
        distance <- distHaversine(p1 = prop_long_lat,
                                  p2 = metro_long_lat)
        output_distances <- c(output_distances, distance)
    }

    return(min(output_distances)/1609.344)
    }
}
```

## Our Wrapper Function
```{r}
distance_function <- function(LONG, LAT, metro_df){
    if(!require(geosphere)){ # what does require do?
      warning("geosphere package is not installed")
    } else{
      # We need to turn our LONG and LAT columns into a vector
      prop_long_lat <- c(LONG, LAT)
      # Now call our prop_metros_dist function from above
      out <- prop_metros_dist(prop_long_lat,
                            metro_df)
      return(out)
    }
}

distance_function(joined_data$LONGITUDE[3],
                  joined_data$LATITUDE[3],
                  metro_data) 
```

## In-Class 8

Combine our function with a loop to find the closest metro station to each property (Hint: Look at one of our previous functions)

```{r, eval = FALSE}
joined_data$metro_distance <- NA #initialize the column
for( )){

  # Your answer here
  
}
```

## Graphing the Relationship

```{r, echo = FALSE, eval = FALSE}
joined_data %>% 
    mutate(distance_tenth = round_any(metro_distance, 0.1))

joined_data %>% group_by(distance_tenth) %>% 
    summarize(price = mean(PRICE/SQUARE.FEET, na.rm = T)) %>% 
    ggplot(aes(x = distance_tenth, y = price)) +
    geom_line() +
    labs(title = "Price vs distance from metro station",
         y = "USD/Sq ft.",
         x = "Miles",
         caption = "Data from Redfin and WMATA")
```