---
title: "Module 1 Day 3"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

```{r}
#install.packages("spatstat")
library(spatstat)
library(tidyverse)
library(ggplot2)
# install.packages("stargazer")
library(stargazer)
# install.packages("Hmisc")
library(Hmisc)
# install.packages("lubridate")
library(lubridate)
```

## Recap last week

Set your working directory

```{r}
setwd("/if/fame/afe/usr/amonken/Howard/Class_Materials/Spring_2018/module_1/day_3")
```

* Read in the data file "acs_2016_sample.csv"
* Calculate the median total income by race, sex, and age
* Make a line plot with age on the x axis and total_income on the y axis
* Use race as the color axis and linetype for sex
* Label your chart appropriately 

```{r}
acs_2016_cleaned <- 
  read.csv("acs_2016_sample.csv") %>%
  rename(state_code = STATEFIP,
         sex = SEX,
         age = AGE,
         race = RACE,
         hispanic = HISPAN,
         education = EDUC,
         employment_status = EMPSTAT,
         total_income = INCTOT,
         wage_income = INCWAGE,
         hrs_worked = UHRSWORK,
         weeks_worked = WKSWORK2,
         weight = PERWT,
         metro = METRO) %>%
  select(state_code, sex, age, race, hispanic, education, 
         employment_status, total_income, wage_income, 
         hrs_worked, weeks_worked, weight, metro) %>%
  filter(total_income <= 1000000,
         wage_income <= 1000000,
         age >= 25,
         age <= 55,
         hrs_worked > 0) %>%
  mutate(sex = ifelse(sex == 1, "Male", "Female"),
         education = case_when(education <= 5 ~ "No HS Diploma",
                               education == 6 ~ "HS Diploma",
                               education %in% c(7,8,9) ~ "Some College",
                               education == 10 ~ "College Degree",
                               education == 11 ~ "Graduate Degree"),
         race = case_when(race == 1 & hispanic == 0 ~ "Non-hispanic White",
                          race == 1 & (hispanic %in% c(1,2,3,4)) ~ "Hispanic White",
                          race == 2 ~ "Black",
                          race == 3 ~ "Native American",
                          race %in% c(4,5,6) ~ "Asian",
                          race %in% c(7,8,9) ~ "Other/Multiracial"))
```

```{r, fig.width=6, fig.height=4}
acs_2016_cleaned %>%
  group_by(age, sex, race) %>%
  summarise(median_income = median(total_income, rm.na = TRUE)) %>%
  ggplot(aes(x = age, y = median_income, color = race, linetype = sex)) +
  geom_line() +
  labs(x = "Age", 
       y = "Income",
       linetype = "Sex",
       color = "Race",
       title = "Median Income by Age, Sex, and Race",
       caption = "Source: 2016 ACS")
```

## facet_wrap()

```{r, fig.width=6, fig.height=4}
acs_2016_cleaned %>%
  group_by(sex, age, race) %>%
  summarise(median_income = median(total_income, rm.na=T)) %>%
  ggplot(aes(x = age, 
             y = median_income, 
             color = sex, 
             linetype = sex)) +
  geom_line() +
  facet_wrap("race") +
  labs(x = "Age", 
       y = "Income",
       linetype = "Sex",
       color = "Sex",
       title = "Median Income by Age, Sex, and Race",
       caption = "Source: 2016 ACS")
```

Why do you think certain lines are noisier than others?


Create a new column in acs_2016_cleaned called rounded age that is the individual age rounded to the nearest 5. (Hint check out the function plyr::round_any()).
Calculate the mean wage income by rounded_age, sex, and education level.
Update the x axis so that it goes from 25-55 in 5 year increments
Update the y axis so that is goes from 0-175,000 in 25,000 increments

```{r}
acs_2016_cleaned %>%
  mutate(rounded_age = plyr::round_any(age,5)) %>%
  group_by(rounded_age, sex, education) %>%
  summarise(mean_wage = mean(wage_income)) %>%
  ggplot(aes(x = rounded_age, y = mean_wage, color = sex)) +
  geom_line() +
  facet_wrap("education", nrow = 3) +
  labs(x = "Age",
       y = "Wage Income",
       title = "Average Wage Income by Education Level and Sex",
       color = "Sex") +
  scale_x_continuous(breaks = seq(25,55,5)) +
  scale_y_continuous(breaks = seq(0,175000,25000))
```

Let's calculate the unweighted and weighted mean income by sex in our acs_2016_cleaned data frame

```{r}
mean_table <-
acs_2016_cleaned %>%
  group_by(sex) %>%
  summarise(unweighted_mean = mean(total_income, na.rm = T),
            weighted_mean = wtd.mean(total_income, weight, na.rm = T))

mean_table
```

## stargazer()

It is very easy to use! It does most of the work for you

```{r}
stargazer(acs_2016_cleaned %>%
          select(age, total_income, 
                 wage_income, hrs_worked, 
                 weeks_worked))
```

If you just run the code it looks like garbage (unless you are fluent in Latex). And it would look like that in the knitted pdf too.

To get it to compile nicely you need to add a chunk option: `results = 'asis'`

```{r, results = 'asis'}
stargazer(acs_2016_cleaned %>%
          select(age, total_income, 
                 wage_income, hrs_worked, 
                 weeks_worked))
```

We can make a table of a data frame by telling stargazer not to make it a summary table. We can turn off the annoying header using `header = FALSE`. We can also set the number of digits to use `digits = 2`

```{r, results = 'asis'}
stargazer(mean_table, summary = FALSE, header = FALSE)
```

We can fix up the column names and show a less ridiculous number of digits

```{r}
mean_table <-
acs_2016_cleaned %>%
  group_by(sex) %>%
  summarise(unweighted_mean = 
              round(mean(total_income, 
                         na.rm = T),0),
            weighted_mean = 
              round(wtd.mean(total_income, weight, 
                             na.rm = T),0))

names(mean_table) <- c("Sex", 
                       "Unweighted Mean", 
                       "Weighted Mean")

```

```{r, results = 'asis'}
stargazer(mean_table, summary = FALSE, header = FALSE, 
          rownames = FALSE,
          title = "Mean Income by Sex")

```

Please make a summary statistics table for the weighted median income by sex

* The weight median function is `weighted.median()`
* Use stargazer to make it beautiful
* Give it an appropriate title

```{r}
median_table <-
acs_2016_cleaned %>%
  group_by(sex) %>%
  summarise(unweighted_median = 
              round(median(total_income,
                         na.rm = T),0),
            weighted_median = 
              round(weighted.median(total_income, weight, 
                             na.rm = T),0))

names(median_table) <- c("Sex", 
                       "Unweighted Median", 
                       "Weighted Median")

stargazer(median_table, summary = FALSE, header = FALSE, 
          rownames = FALSE,
          title = "Median Income by Sex")
```



Let's make a quick chart of the distribution of income by gender and then make if fabulous

* What geom should you use? What are the aesthetics it needs?
* To start make a density chart of total income by gender. 
* Be sure to title and label your chart appropriately

```{r, echo = FALSE, fig.width=6, fig.height=4}
inc_sex_plot <-
acs_2016_cleaned %>%
  ggplot(aes(x = total_income, fill = sex, color = sex)) +
  geom_density(alpha = 0.5) +
  labs(x = "Total Income",
       y = "Density",
       fill = "Sex",
       color = "Sex",
       title = "Distribution of Income by Sex",
       caption = "Source: 2016 ACS")

inc_sex_plot
```

We can use the scale_color_manual function to change the colors. 
```{r, fig.width=6, fig.height=3}
inc_sex_plot + 
  scale_color_manual(values = c("dodger blue", "orange"))+ 
  scale_fill_manual(values = c("dodger blue", "orange"))
```


Make a fancy chart of the distribution of total income by education level
* label your chart appropriately
* use a non-base theme and make things colorful
* Here's a link for many named colors in R http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf


```{r}
acs_2016_cleaned %>%
  ggplot(aes(x = total_income, fill = education, color = education)) +
  geom_density(alpha = 0.3) +
  labs(x = "Total Income",
       y = "Density",
       fill = "Education",
       color = "Education",
       title = "Distribution of Income by Eduaction",
       caption = "Source: 2016 ACS") + 
  scale_color_manual(values = c("dodger blue", "purple","firebrick","goldenrod","palegreen3"))+ 
  scale_fill_manual(values = c("dodger blue", "purple","firebrick","goldenrod","palegreen3"))

```


## unique()

## Urban vs. rural

I want to know the average total income by race for people in metro areas and outside of metro areas (rural)

## unique()

What are the values of the metro variable?

* We can find out using the `unique()` function
* But what does it do?

```{r}
unique(c("A", "B", "A", "B", "C", "C", "D"))
```

## unique()

How many categories of metro status are there?

```{r}
unique(acs_2016_cleaned$metro)
```

## Urban vs. rural

I want to know the average total income by race for people in metro areas and outside of metro areas (rural)

* First filter to people we know for sure are not in a metro area or in the center or outskirts of a metro area. 
    + 0		Not identifiable
    + 4		Central / Principal city status unknown

* Then recode the metro area 
    + 1	~ "Rural"
    + 2	~ "Central / Principal City"
    + 3	~ "Outside Central / Principal City"

```{r}
acs_2016_cleaned <-
  acs_2016_cleaned %>%
  filter(!(metro %in% c(0,4))) %>%
  mutate(metro = case_when(metro == 1 ~ "Rural",
                           metro == 2 ~ "Central / Principal City",
                           metro == 3 ~ "Outside Central / Principal City"))
```


* Now calculate the average total income by race and metro status
* Be sure to take the weighted mean
* Make a faceted bar chart with race on the x axis and mean income on the y axis
* Use metro as the faceting variable
* Be sure to pick a theme and use scale_x_ to keep the labels from overlapping

```{r, echo = FALSE, fig.width=6, fig.height=4}
acs_2016_cleaned %>%
  group_by(race, metro) %>%
  summarise(mean_income = wtd.mean(total_income, weight, na.rm = F)) %>%
  ggplot(aes(x = race, y = mean_income, color = race, fill = race)) +
  geom_col(alpha = 0.5) + 
  facet_wrap("metro", ncol = 1) + 
  labs(x = "Race",
       y = "Income",
       title = "Mean Income by Race and Metropolitan Area",
       color = "Race",
       fill = "Race") +
  theme_minimal() +
  scale_x_discrete(labels= c("Asian", "Black", "Hispanic \nWhite", "Native \nAmerican", "Non-hispanic \nWhite", "Other/\nMultiracial")) +
  guides(color = FALSE,
         fill = FALSE)
```

* Make the same chart, faceting on race.

```{r, echo = FALSE, fig.width=6, fig.height=4}
acs_2016_cleaned %>%
  group_by(metro, race) %>%
  summarise(mean_income = wtd.mean(total_income, weight, na.rm = F)) %>%
  ggplot(aes(x = metro, y = mean_income, color = race, fill = race)) +
  geom_col(alpha = 0.5) + 
  facet_wrap("race", ncol = 3) + 
  labs(x = "Metro",
       y = "Income",
       title = "Mean Income by Metropolitan Area and Race",
       color = "Race",
       fill = "Race") +
  theme_minimal() +
  scale_x_discrete(labels= c("Asian", "Black", "Hispanic \nWhite", "Native \nAmerican", "Non-hispanic \nWhite", "Other/\nMultiracial")) +
  guides(color = FALSE,
         fill = FALSE)
```

## Dates

We're going to learn how to make this chart.

```{r, fig.width=4.5, fig.height=4.5*0.666, fig.align='center',echo=FALSE}
fred_median_earnings <- 
  read.csv("fred_median_earnings.csv", 
           stringsAsFactors = F)
fred_median_earnings %>%
  mutate(Date = as.Date(Date, format("%m/%d/%Y"))) %>%
  ggplot(aes(x = Date, y = Median_earnings)) +
  geom_line() +
  labs(x = "Date", 
       y = "Weekly Earnings",
       title = "Real Median Weekly Earnings over Time",
       caption = "Source: FRED") +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y") +
  theme_minimal()
```

## Dates

* Dates are very special in R
* R understands dates of the form:

```{r}
as.Date("2012-08-30")
```
* We can do standard arithmetic on the dates:

```{r}
as.Date("2012-08-30") - 3
```
* What happens if we subtract two dates from each other?

## Dates

But what if our date is not in that format? What if we are importing data?

```{r}
as.Date("08/30/2012", format = "%m/%d/%Y")
as.Date(41149, origin = "1900-01-01")
```

* %d -> Day
* %m -> Numeric Month 
* %b -> Abbreviated Month
* %B -> Full Month
* %y -> 2-digit year
* %Y -> 4-digit year

## Dates

Convert the following to date objects

* "Jan 1, 2018" 
* "31/1/18"
```{r}
as.Date("Jan 1, 2018", format = "%b %d, %Y")
as.Date("31/1/18", format = "%d/%m/%y")

```

## Dates

We can also make sequences of dates
```{r}
# First day of every month
seq(as.Date("2018-01-01"), 
    as.Date("2018-04-01"), 
    by = "months")
```
```{r}
seq(as.Date("2018-01-01"), 
    as.Date("2018-01-05"), 
    by = "2 days")
```

## Dates

We can also subtract from sequences of dates!
```{r}
# Last day of every month
seq(as.Date("2018-01-01"), 
    as.Date("2018-04-01"), 
    by = "months") - 1
```

## Dates

Let's look at time series of real median weekly earnings:

* Read in the file fred_median_income.csv
* What columns are in the dataframe?

```{r}
fred_median_earnings <- 
  read.csv("fred_median_earnings.csv", 
           stringsAsFactors = F)
```

## Dates

* Convert the date column to a date object
* Make a line chart with the date on the x axis and Median_earnings on the y axis
* Be sure to label your chart correctly
* Use the scale_x_dates() function to label every 2 years

## Dates

```{r, fig.width=6, fig.height=4, echo=FALSE}
fred_median_earnings %>%
  mutate(Date = as.Date(Date, format("%m/%d/%Y"))) %>%
  ggplot(aes(x = Date, y = Median_earnings)) +
  geom_line() +
  labs(x = "Date", 
       y = "Weekly Earnings",
       title = "Real Median Weekly Earnings over Time",
       caption = "Source: FRED") +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y") +
  theme_minimal()
```

* How have average weekly earnings recovered since the great recession? How have they changed since 2000?

## Dates: `lubridate`

* As you can see, we are currently working with quarterly data. 
* Let's convert this to annual data using the package `lubridate`. 
* `lubridate` makes it easy to extract parts of a date from date objects.

## Dates: `lubridate`

* In this case, we will need to extract the year from our date column.

```{r}
library(lubridate)
date <- as.Date("2018-02-09")

# What is the year of this date?
year(date)

# What is the month of this date?
month(date)
```

## Dates: `lubridate`

* Create a new column of years
* Use `group_by()` and `summarise()` to create annual averages of median earnings
* Create a line chart exactly like the one before
* HINT: you won't be able to use `scale_x_date()` on the year variable

## Dates: `lubridate`

```{r, fig.width=6, fig.height=4, echo=FALSE}
fred_median_earnings %>%
  mutate(Date = as.Date(Date, format("%m/%d/%Y")),
         year = year(Date)) %>%
  group_by(year) %>%
  summarise(Median_earnings = mean(Median_earnings)) %>%
  ggplot(aes(x = year, y = Median_earnings)) +
  geom_line() +
  labs(x = "Date", 
       y = "Weekly Earnings",
       title = "Real Median Weekly Earnings over Time",
       caption = "Source: FRED") +
  scale_x_continuous(breaks = seq(2000, 2017, 2)) +
  theme_minimal()
```

* When might it be useful to aggregate data to a lower frequency?

## Housekeeping

* REMINDER: your first project assignment is due next week (Friday, 2/16).
* This is your **project proposal** document.

## Housekeeping

* Homework for the entire module is in one document, **Module_1_STUDENT.Rmd**.
* Look under the "Day 1", "Day 2", and "Day 3" headers.
* After today, you will work on the "Day 3" portion of the document.